```{r create_basic_documentation, echo=TRUE}
doc <- NULL
doc$run.date <- date()
doc$version <- system(' git rev-parse HEAD', intern=TRUE)
doc$author <- "Steve Simon (KUMC)"
doc$maintainer <- "Steve Simon (KUMC)"
doc$assistants <- "Dan Connolly"
```


Case-control
============
For context, see [485].

[485]: https://informatics.gpcnetwork.org/trac/Project/ticket/485
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

This program analyzes a case-control data set. It produces some simple
tables and graphs based on the lasso and/or elastic net models.

A companion program, extract_case_control, produces the data set
used by this program.

This program was run on `r doc$date` using version `r doc$version`.
The original author is `r doc$author`. `r doc$maintainer`
is currently maintaining and enhancing this program
with the assistance of `r doc$assistants`.


```{r load_required_libraries, echo=TRUE}
save.image("backup.RData")
rm(list=ls())
cat("\n\nQuality check: Are we in the correct directory?")
getwd()
# Don't wrap so much
options(width=90)
# Load the case-control data sets
load("case_control_data.RData")
start_time <- Sys.time()
print(start_time)
display_elapsed_time <- function() {
  cat("\n\nThis code chunk required ")
  print(round(Sys.time()-start_time))
  return(Sys.time())
}
```

Here's a summary of how the data was selected. We
identified four disease groups in Heron (KUMC's i2b2).

Breast cancer (as identified in the SEER Site Summary),

All other cancers (again from SEER),

Diabetes (ICD9 250 or ICD10 E08-E13), and

Ischemic heart disease (ICD9 410-414.99 or ICD10 I20-I25).

All males were excluded.

The first group represents cases and the other three
represent different control groups. Patients who qualified
for two or more of the above groups were excluded from the
analysis.

There were no date restrictions. In future versions, we may want
to consider some date restrictions and try matching subjects on
age.

We then merged that data with the PCORNET CDM to get 
any ICD9 codes associated with each patient.

Below, we include only those codes that appear reasonably often
among the cases.

```{r include only the popular icd9 codes, echo=TRUE}
cutoff <- 100
n1 <- length(unique(lb$DX))
lb <- lb[lb$dx_count >= cutoff, ]
n2 <- length(unique(lb$DX))
start_time <- display_elapsed_time()
```

The initial data set has `r n1` unique ICD9 codes. After eliminating
the codes that occur less than `r cutoff` times among the cases,
there are `r n2` unique ICD9 codes left.

```{r experiment_with_sparse_matrices, echo=TRUE}
library("Matrix")
start_time <- display_elapsed_time()
i <- i.control[1]
cc1 <- lb$GP %in% c("breastcancer", i)
tst <- lb[cc1, ]
i <- factor(tst$PATIENT_NUM)
j <- factor(tst$dx_label)
sm <- sparseMatrix(i=as.numeric(i), j=as.numeric(j), x=1)
sm[1:90,1:2]
dim(sm)
start_time <- display_elapsed_time()
cc1 <- lb$GP %in% c("breastcancer", i)
pc1 <- dcast(lb[cc1, ],PATIENT_NUM + GP ~ dx_label,length)
pc1[1:90,3:4]
dim(pc1)
start_time <- display_elapsed_time()
```

```{r store_lasso_models, echo=TRUE}
library("reshape2")
library("glmnet")
all_models <- NULL
all_cv <- NULL
for (i in i.control) {
  cc1 <- lb$GP %in% c("breastcancer", i)
  pc1 <- dcast(lb[cc1, ],PATIENT_NUM + GP ~ dx_label,length)
  en <- glmnet(
    x=as.matrix(pc1[,-(1:2)]), alpha=0.5, 
    y=as.numeric(pc1$GP=="breastcancer"),
    family="binomial")
  all_models[[i]] <- en
  cv <- cv.glmnet(
    x=as.matrix(pc1[,-(1:2)]), alpha=0.5, 
    y=as.numeric(pc1$GP=="breastcancer"),
    lambda=en$lambda,
    family="binomial",
    type.measure="class")
  all_cv[[i]] <- cv
}
start_time <- display_elapsed_time()
```

```{r display_lasso_models, echo=TRUE, fig.width=10, fig.height=12}
best_variables <- NULL
for (ic in i.control) {
  cv <- all_cv[[ic]]
  plot(cv)

  en <- all_models[[ic]]
  n_lambda <- dim(en$beta)[2]
  n_variables <- dim(en$beta)[1]
  lambda_range <- log(range(en$lambda))
  entry_point <- apply(en$beta,1,function(x) sum(abs(x)==0))
  x_position <- log(en$lambda)[entry_point]
  coefficient_direction <- (1+sign(en$beta[,n_lambda]))/2
  step <- which(cv$lambda==cv$lambda.1se)
  title(paste("CV suggests using step", step), cex=0.5)
  best_variables[[ic]] <- names(entry_point)[entry_point <= step]

  for (d in 1:0) {
    direction_label <- c("negative","positive")[d+1]
    plot(0, 0,
      xlim=c(-1,1)*max(abs(en$beta)), 
      ylim=c(log(cv$lambda.1se),log(max(en$lambda))),
      type="n",
      xlab="Beta",
      ylab="log(lambda)")
    title1 <- paste("Lasso/elastic net model comparing",i.case,"to",ic)
    title2 <- paste(direction_label,"coefficients only.")
    title(paste(title1,title2,sep="\n"))
    for (i in 1:n_variables) {
      if (coefficient_direction[i]==1-d) next
      lines(en$beta[i,1:step],log(en$lambda))
    }
    cat("\n\nVariables influencing", i.case, "vs", ic)
    cat(" in a", direction_label,"direction.")
    cat("\nRanked from most important to least important.\n")
    for (i in 1:step) {
      variable_subset <- entry_point==i & coefficient_direction==d
      variable_list <- names(entry_point)[variable_subset]
      variable_label <- paste(variable_list, collapse="\n")
      if (sum(variable_subset) > 0) {
        cat("\nStep", i, "\n")
        cat(paste("  ", variable_list, "\n", sep=""), sep="")
      }
      text(0.1-0.2*d, log(en$lambda[i]), variable_label, adj=d)
    }
  }
}
consensus_variables <- best_variables[[1]]
for (i in 2:length(best_variables)) {
  consensus_variables <- intersect(consensus_variables, best_variables[[i]])
}
cat("\n\nHere is a list of variables found in all three models")
print(consensus_variables)
start_time <- display_elapsed_time()
```

```{r save-everything, echo=TRUE}
save.image(file="analyze_lasso.RData")
start_time <- display_elapsed_time()
```
