```{r create-basic-documentation, echo=TRUE}
doc <- NULL
doc$run.date <- date()
doc$version <- system(' git rev-parse HEAD', intern=TRUE)
doc$author <- "Steve Simon (KUMC)"
doc$maintainer <- "Steve Simon (KUMC)"
doc$assistants <- "Dan Connolly"
```

This program reads data from an i2b2 query and produces
simple analyses and graphs. It stores the data and some
intermediate files in an RData file.

This program was run on `r doc$date` using version `r doc$version`.
The original author is `r doc$author`. `r doc$maintainer`
is currently maintaining and enhancing this program
with the assistance of `r doc$assistants`.

```{r load-required-libraries, echo=TRUE}
library("ggplot2")
library("reshape")
library("RSQLite")
save.image("backup.RData")
rm(list=ls())
```

```{r set-options, echo=TRUE}

# Don't wrap so much
options(width=100)

```

Case-control
============
For context, see [485].

[485]: https://informatics.gpcnetwork.org/trac/Project/ticket/485
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

```{r setup-sql-queries, echo=TRUE}

# use-this-database.txt should have one entry for each file.
# The first file is presumed to be the cases and each of the 
# remaining files is presumed to be a different control group.

setwd("/d1/home/ssimon/bc_qa/bc_case_control")
file_info <- read.csv(file="use_this_database.txt", 
                      header=TRUE, 
                      stringsAsFactors=FALSE)
print(file_info)
variable_groups <- file_info$variable_group

sql_query <-NULL

# There are two ways to query the data. The first way
# is intuitive (to me) but inefficient. The second way
# is less intuitive but more efficient because it relies
# on SQL rather than R to count.

# count the overall number of patients in each disease group.
sql_query[["disease_group_counts"]] <-
  "select count(distinct patient_num) AS n
   from observation_fact"

# get distinct patient numbers in each disease group.
sql_query[["distinct_patients"]] <-
  "select distinct patient_num
   from observation_fact"

# read the observation_fact table, removing duplicates.
sql_query[["patient_level_concepts"]] <-
  "select distinct obs.patient_num, obs.concept_cd
   from observation_fact obs"

# count the number of patients who have each fact.
sql_query[["concept_counts"]] <-
  "select concept_cd, count(distinct patient_num) AS n
   from observation_fact
   group by concept_cd"


```

```{r get-overall-patient-counts}
# dp: distinct patient numbers in each disease group

g0 <- NULL
dp <- NULL
file_names <- file_info$file_location[variable_groups=="patients"]
group_names <- file_info$disease_group[variable_groups=="patients"]
for (i in 1:length(file_names)) {
  file_name <- file_names[i]
  group_name <- group_names[i]
  cat("\nRunning queries for ", group_name, " (", file_name,  ").\n", sep="")
  conn_site <- dbConnect(SQLite(), dbname=file_name)
  g0[[group_name]] <- dbGetQuery(conn_site, sql_query["concept_counts"])
  cat("\n\nThis is the number of concepts used to define",group_name,".\n\n")
  print(dim(g0[[group_name]][1]))
  cat("\n\nHere are the first few concepts used to define",group_name,".\n\n")
  print(head(g0[[group_name]], 20))
  dp[[group_name]] <- dbGetQuery(conn_site, sql_query["distinct_patients"])
  cat("\n\nThis is the number of patients in",group_name,".\n\n")
  print(dim(dp[[group_name]][1]))
}
```

```{r list-patient_selection-codes}
selection_codes <- g0[[1]]$concept_cd
for (i in 2:length(file_names)) {
  selection_codes <- union(selection_codes, g0[[i]]$concept_cd)
}
print(sample(selection_codes,100))
```

```{r get-event-counts}
g1 <- NULL
g2 <- NULL
file_names <- file_info$file_location[variable_groups=="procedure"]
group_names <- file_info$disease_group[variable_groups=="procedure"]
for (i in 1:length(file_names)) {
  file_name <- file_names[i]
  group_name <- group_names[i]
  cat("\nRunning queries for ", group_name, " (", file_name,  ").\n", sep="")
  conn_site <- dbConnect(SQLite(), dbname=file_name)
  cat("\nQuality check: compare this...\n")
  g1[[group_name]] <- dbGetQuery(conn_site, sql_query["patient_level_concepts"])
  print(head(table(g1[[group_name]]$concept_cd)))
  g2[[group_name]] <- dbGetQuery(conn_site, sql_query["concept_counts"])
  cat("\nto this...\n\n")
  print(head(g2[[group_name]]))
  dbDisconnect(conn_site)
}
sapply(g1,dim)
sapply(g2,dim)
cc <- g2[[1]]
names(cc)[2] <- "n1"
for (i in 2:length(group_names)) {
  cat("\nMerging ", group_names[i], ".\n\n", sep="")
  cc <- merge(cc, g2[[group_names[i]]], by="concept_cd", all=TRUE)
  names(cc)[i+1] <- paste("n",i,sep="")
}
cc[is.na(cc)] <- 0
head(cc,100)
tail(cc,100)

```

```{r remove-selection_codes}
# There are hundreds if not thousands of codes used to select the various
# disease groups. You need to remove these before doing any serious
# data analysis.
cc <- merge(cc,data.frame(concept_cd=selection_codes,exclude=1),all=TRUE)
cc$exclude[is.na(cc$exclude)] <- 0
cc <- cc[cc$exclude==0,]
head(cc,50)
tail(cc,50)
```


```{r reshape-event-information, echo=TRUE, eval=FALSE}

library("reshape2")

# equal to 1 unless NULL

db$all <- rbind(data.frame(db$ca,group="t+;d+"), data.frame(db$co,group="t+;d-"))

db$counts <- dcast(db$all, group ~ concept_cd, length)
db$counts[, 1] <- as.character(db$counts[, 1])
db$counts[1:2, 1:5]

unique_concepts <- db$concepts[!duplicated(db$concepts$concept_cd), ]
predictors <- unlist(dimnames(db$counts)[2])[-1]
labels <- merge(data.frame(concept_cd=predictors),unique_concepts)
labels$name_char <- substr(labels$name_char,1,50)
head(labels)
# unlist will turn the 1 by n matrix into a vector.
t_pos_d_pos  <- unlist(db$counts[1,-1])
t_pos_d_neg <- unlist(db$counts[2,-1])

t_neg_d_pos <- n_ca - t_pos_d_pos
t_neg_d_neg <- n_co - t_pos_d_neg

ppv  <- round(100*t_pos_d_pos / (t_pos_d_pos+t_pos_d_neg))
npv  <- round(100*t_neg_d_neg / (t_neg_d_pos+t_pos_d_neg))

sens <- round(100*t_pos_d_pos / (t_pos_d_pos+t_neg_d_pos))
spec <- round(100*t_neg_d_neg / (t_neg_d_neg+t_pos_d_neg))

lp <- "("
rp <- ")"
sl <- "/"
sp <- " "
pc <- "%"
ppv_labels <- paste(ppv, pc, sp, lp, t_pos_d_pos, 
                    sl, t_pos_d_pos+t_pos_d_neg, rp,
                    sp,predictors,sp,labels$name_char,sep="")
high_ppv <- ppv>90 & t_pos_d_pos > 100
# Re-arrange so that highest numerators appear first.
o <- rev(order(t_pos_d_pos[high_ppv]))
ppv_labels[high_ppv][o]
```


```{r save-everything, echo=TRUE}
save.image(file="case_control.RData")
```
