```{r create_basic_documentation, echo=TRUE}
# Before anything else, set global options
opts_chunk$set(echo=TRUE, cache=TRUE, error=TRUE)

doc <- NULL
doc$run.date <- date()
doc$version <- system(' git rev-parse HEAD', intern=TRUE)
doc$author <- "Steve Simon (KUMC)"
doc$maintainer <- "Steve Simon (KUMC)"
doc$assistants <- "Dan Connolly"
```


Case-control
============
For context, see [485].

[485]: https://informatics.gpcnetwork.org/trac/Project/ticket/485
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

This program takes output from the lasso and/or elastic net models
and produces some simple tables and graphs.

A companion program, analyze_lasso, produces the lasso or elastic net
models used by this program and analyze_lasso, in turn, relies on 
extract_case_control which produces the data sets needed.

This program was run on `r doc$date` using version `r doc$version`.
The original author is `r doc$author`. `r doc$maintainer`
is currently maintaining and enhancing this program
with the assistance of `r doc$assistants`.


```{r load_required_libraries, echo=TRUE}

save.image("backup.RData")
rm(list=ls())

verbose=TRUE # if TRUE, print various intermediate values and quality checks.

if (verbose) {
  cat("\n\nQuality check: Are we in the correct directory? ")
  print(getwd())
}

# Don't wrap so much
options(width=90)

# load the required libraries
library("reshape2")
library("glmnet")
library("Matrix")

# Load the case-control data sets
load("analyze_lasso.RData")

rm(timing_log)
start_time <- Sys.time()
if (verbose) {
  print(start_time)
}

# Here are some functions needed in this program.

source("create_utility_functions.R")

```

A summary of how the data were originally collected appears in 
extract_case_control.Rmd.

The lasso and elastic net models have a tuning parameter, lambda,
that controls the complexity of the model. You can pick the
optimal value of lambda through cross validation.

```{r find_optimal_lambda, timer=TRUE}
optima <- NULL
for (ic in i.control) {
  cv <- all_cv[[ic]]
  optima[[ic]]$lambda <- cv$lambda.1se # maybe try lambda.min
  optima[[ic]]$step <- which(cv$lambda==cv$lambda.1se)
  cat("\nComparison of", i.case, "to", ic)
  cat(".\n  Best value of lambda is", optima[[ic]]$lambda)
  cat(", which is step", optima[[ic]]$step,"\n")
}
```

Now that you have the optimum, pull out the coefficients for that
particular value of lambda. You should distinguish between coefficients
that are positive (direction=1), negative (direction=0), or zero
(direction=0.5). Positive coefficients are ones that increase the 
estimated probability of being a case and negative coefficients 
are ones that increase the estimated probability of being a 
control. One of the attractions of the lasso/elastic net is
that it zeros out the coefficients of many variables, and thus
it serves as a feature selection method.

The other big advantage of the lasso is that it shrinks all 
non-zero coefficients towards zero. Some are shrunk more than others
but the process of shrinkage helps avoid overfitting.
towards zero, which helps 

```{r extract_coefficients_at_optimum, timer=TRUE}
coefficients <- NULL
for (ic in i.control) {
  op <- optima[[ic]]
  en <- all_models[[ic]]
  direction <- (1+sign(en$beta[,op$step]))/2
  # Note the following statement will fail if a coefficent starts out in
  # a certain direction and then returns to zero. This is rare, but it
  # could happen. Maybe use min(which(x!=0)) instead.
  entry_step <- apply(en$beta,1,function(x) sum(abs(x)==0))
  entry_lambda <- en$lambda[entry_step]
  beta <- en$beta[,op$step]
  odds_ratio <- round(exp(beta),2-direction)
  coefficients[[ic]] <- 
    data.frame(odds_ratio, beta, direction, entry_step, entry_lambda)
  if (verbose) {
    cat("\nComparison of", i.case, "to", ic,"\n\n")
    print(list_random_rows(coefficients[[ic]]))
  }
}
```

The following graphs show the development of the lasso/elastic
net models from the simplest models to model indicated by
the optimal lambda. The graphs are split out by positive versus
negative coefficients in part to simplify the presentation. But
also, the terms associated with positive coefficients appear
to have better consistency across multiple control groups than
terms associated with negative coefficients.

Notice that variables that enter the model early are often,
but not always, ones with large coefficients (large meaning
far away from the null value of 1.0) in the optimal
model. Also notice that the positive terms enter earlier and
end up at more extreme odds ratios than the negative terms.

```{r draw_lasso_graphs, timer=TRUE}
getwd()
positive_only <- 1
positive_and_negative <- 1:0
bmp(file="dl%02d.bmp", width=800, height=800)
for (ic in i.control) {
  en <- all_models[[ic]]
  op <- optima[[ic]]
  co <- coefficients[[ic]]
  for (d in positive_only) {
    direction_label <- c("Negative", "Positive")[d+1]
    yl <- c(0, 1)
    if (d==0) {yl <- c(-1, 0)}
    plot(0, 0,
      ylim=yl*max(abs(co$beta)), 
      xlim=c(log(max(en$lambda)),log(op$lambda)*1.5),
      type="n",axes=FALSE,
      ylab="Odds ratio",
      xlab="log(lambda)")
    title1 <- paste("Lasso/elastic net model comparing",i.case,"to",ic)
    title2 <- paste(direction_label,"coefficients only.")
    title(paste(title1,title2,sep="\n"))
    axis(side=1)
    or_labels <- c(1,2,5,10,20,50,100)
    if (d==0) {or_labels <- 1/or_labels}
    axis(side=2, at=log(or_labels), labels=or_labels)
    # find all coefficients going in one direction.
    i.direction <- which(co$direction == d)
    for (id in i.direction) {
      lines(log(en$lambda[1:op$step]), en$beta[id,1:op$step])
      text(log(op$lambda), co$beta[id], dimnames(co)[[1]][id], adj=0)
    }
  }
}
dev.off()
```

The consensus across the models will involve selecting
coefficients that are consistent in direction across
multiple control groups.

```{r develop_consensus, timer=TRUE}
average_beta <- rep(0,dim(coefficients[[1]])[1])
cumulative_direction <- rep(0,dim(coefficients[[1]])[1])
for (ic in i.control) {
  average_beta <- average_beta + coefficients[[ic]]$beta
  cumulative_direction <- cumulative_direction + 2*coefficients[[ic]]$direction - 1
}
average_beta <- average_beta / length(i.control)
or_consensus <- data.frame(
  lb=dimnames(coefficients[[1]])[[1]],
  gm=round(exp(average_beta),1),
  co=round(average_beta,2),
  di=cumulative_direction)
for (i in 1:length(i.control)) {
  or_consensus <- data.frame(or_consensus, coefficients[[i.control[i]]]$odds_ratio)
  names(or_consensus)[i+4] <- i.control[i]
}
o <- rev(order(average_beta))
or_consensus <- or_consensus[o, ]
# Odds ratios for consensus positive coefficients.
or_consensus[or_consensus$di==3, c("lb", "gm", "co")]
# Individual odds ratios for consensus.
or_consensus[or_consensus$di==3, i.control]
# These variables did not make the cut. The column
# for di indicates how many models (0, 1, or 2)
# included this variable. It looks like a fair
# number of promising variables did not make
# the cut.
or_consensus[or_consensus$di < 3, c("di", "lb")]
```


```{r save-everything, timer=TRUE}
save.image(file="display_lasso.RData")
```

Well done. Here is how long everything took.

```{r display_timing_log}
if (verbose) {
  cat("Program began at ")
  cat(as.character(start_time))
  cat("\nProgram ended at ")
  cat(as.character(Sys.time()))
  cat("\n\n")
  tm <- read.table("timing_log.txt")$V1
  cat(paste(tm, collapse="\n"))
}
```

