```{r create_basic_documentation, echo=TRUE}

# As I am learning git, I want to put a few comments in to help me
# track which files appear where. I'll remove these once I get
# more comfortable with git.

# This is a test change made at 11:12am on Monday, June 27.


# Before anything else, set global options
opts_chunk$set(echo=TRUE, cache=FALSE, error=TRUE)

doc <- NULL
doc$run.date <- date()
doc$version <- system(' git rev-parse HEAD', intern=TRUE)
doc$author <- "Steve Simon (KUMC)"
doc$maintainer <- "Steve Simon (KUMC)"
doc$assistants <- "Dan Connolly"
```

First things first. If you run any of these programs, they 
will near the start use a command rm(list=ls()) that cleans
out any files left from previous programs. This helps improve
the reproducibility of the programs. But if you are in the 
habit of storing information in the default location,
.RData, this program may accidentally wipe out your stuff.

I've included a save.image("backup.Rdata") command but that
only provides a fig leaf of protection. You should run these
programs in a fresh directory or backup your .RData file.

evaluate_model.Rmd
==================

For context, see [485].

[485]: https://informatics.gpcnetwork.org/trac/Project/ticket/485
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

This program takes a model and evaluates it on patients. It stores the data and some
intermediate files in an RData file.


Currently, this program takes the model developed by display_lasso.Rmd,
but it can be adapted for models from other programs without too much
difficulty.

Here's a rough outline of how the program works.

1. Get a list of all the variables included in the model. 
2. Get a list of patients.
3. Find any records for these patients that have variables in the model.
4. Calculate a prediction based on those variables.
5. Assess the quality of that prediction.

It requires the following libraries.

1. chron
2. knitr
3. ROracle
4. RSQLite

This program was run on `r doc$date` using version `r doc$version`.
The original author is `r doc$author`. `r doc$maintainer`
is currently maintaining and enhancing this program
with the assistance of `r doc$assistants`.


```{r run_preliminaries, cache=FALSE}

# Backup image (just in case) and then start with a blank slate.

save.image("backup.RData")
rm(list=ls())

# Set qc option, if TRUE, print various intermediate values and quality checks.

qc <- TRUE

# Document when this program started.

start_time <- Sys.time()
if (qc) {
  cat("\n\nProgram started at ")
  print(start_time)
}

# load required libraries.

library("chron")
library("knitr")
library("ROracle")
library("RSQLite")

# Check to see if you are in the proper subdirectory.

if (qc) {
  cat("\n\nQuality check: Are we in the correct directory?\n")
  print(getwd())
}

# Control wrapping

options(width=90)

# Here are some functions needed in this program.

source("create_utility_functions.R")

```

```{r load_data_from_previous_programs, timer=TRUE}
load(file="case_control_data.RData")
load(file="lasso_coefficients.RData")
load(file="display_lasso.RData")

if (exists("arc")) rm(arc)

```

Read the model. Get code_key to link to original names.

Note: the data frames have short names by necessity, but this
makes things a bit cryptic. The data frame lc has all the data,
while ld has the training data only and le has the test data only.

```{r read_model, timer=TRUE}
ma <- merge(x=lc, y=lasso_coefficients, all=FALSE, by.x="dx_label", by.y="dx")
if (qc) {
  print_random_rows(ma)
}
archive(ma, "data with only significant predictors")
```

This is an inner join (all=FALSE), so will exclude any patients who have not
experienced any of the events selected by the lasso model. Fix this by re-merging
with pc.

This code is poorly written because there is a gp variable from mx and a GP
variable from pc. I'll fix this later.

```{r include_excluded_patients, timer=TRUE}
mx <- merge(ma, pc, by.x="patient_num", by.y="PATIENT_NUM", all=TRUE)
mx$co[is.na(mx$co)] <- 0
if (qc) {
  print(summary(mx$co))
  print(table(pc$GP))
  print(table(mx$GP[mx$co==0]))
}
archive(mx, "include excluded patients")
```

For some applications, you want to see WHEN the various events occurred first.
This would be a whole lot easier with the duplicated function, but duplicated
is very inefficient for large data sets.

Note: keep the GP variable from PC rather than the gp variable from ma.

```{r remove_second_and_later, timer=TRUE}
# order by patient, dx, date
o <- order(mx$patient_num, mx$dx_label, mx$observation_date)
mb <- mx[o, ]
mb$dx_label[is.na(mb$dx_label)] <- "Nothing"
n <- dim(mb)[1]
mb$same_patient <- c(0,as.numeric(mb$patient_num[-1]==mb$patient_num[-n]))
mb$same_dx_label <- c(0,as.numeric(mb$dx_label[-1]==mb$dx_label[-n]))
mb$duplicates <- pmin(mb$same_patient, mb$same_dx_label)
# Here's the logic: the first observation is never the same patient or same dx.
# If the patient or dx is unchanged between a given row mb[-1] and the previous
# row mb[-n], then the row is the same patient or same dx.
# A duplicate occurs only if patient is unchanged (same_patient==1) 
# AND dx is unchanged (same_dx==1).
if (qc) print_random_rows(mb)
archive(mb, "remove second and later dx")
keep_list <- c("dx_label","patient_num","GP","dx_count","observation_date","co","TUMOR_DATE")
mc <- mb[!mb$duplicates, keep_list]
# order by patient, date
o <- order(mc$patient_num, mc$observation_date)
mc <- mc[o, ]
archive(mc, "re-ordered")
```

```{r calculate_predictions_over_time, timer=TRUE}
update_probability <- function(baseline_probability, co) {
  baseline_odds <- baseline_probability / (1-baseline_probability)
  updated_odds <- baseline_odds * exp(co)
  return(updated_odds / (1+updated_odds))
}

md <- aggregate(mc$co, mc[, c("patient_num", "GP")], sum)
names(md)[3] <- c("co")
md$probability <- update_probability(0.1, md$co)
if (qc) {
  print_random_rows(md)
  table(md$GP)
}
archive(md, "aggregated to a single patient")
plot(factor(md$GP), md$probability)
tapply(md$probability,md$GP, function(x) {round(sum(x>0.9)/length(x>0.9), 2)})
tapply(md$probability,md$GP, function(x) {round(sum(x==0.1)/length(x==0.1), 2)})

```

Now let's draw some ROC curves.

```{r draw_roc, timer=TRUE}
# the following statement is only needed until we get the data restricted databases.
mg <- split(md$probability, md$GP)
probability_list <- sort(unique(md$probability))
n_p <- length(probability_list)
mh <- matrix(NA, nrow=n_p+2, ncol=length(unique(md$GP)))
dimnames(mh)[[2]] <- sort(c(i_case, i_control))
mh[1, ] <- 1
mh[n_p+2, ] <- 0
# names(mh) <- sort(c(i_case, i_control))
for (i in 1:length(probability_list)) {
  p <- probability_list[i]
  mh[i+1, ] <- sapply(mg, function(x) {sum(x > p) / length(x)})
}
for (ic in i_control) {
  plot(mh[, ic], mh[, i_case], xlim=c(0, 1), ylim=c(0, 1), type="l")
  title(paste("ROC curve for", i_case, "vs", ic))
}
```

Now let's peek at random cases.

The code is a bit of a mess because the format for observation_date is not
the same as the format for TUMOR_DATE.

```{r peek_at_random, fig.width=8, fig.height=5, eval=TRUE}
calculate_time_course <- function(p) {
  sb <- mc[mc$patient_num==p, ]
  sb$cumulative_co <- cumsum(sb$co)
  sb$cumulative_probability <- update_probability(0.1, sb$cumulative_co)
  sb$days <- round(as.numeric(sb$observation_date - sb$observation_date[1]) / (24*60*60), 1)
  sb$days[is.na(sb$days)] <- 0
  sb$tumor_day <- as.numeric(sb$TUMOR_DATE - as.Date(sb$observation_date[1]))
  return(sb)
}
plot_time_course <- function(sb) {  
  n <- dim(sb)[1]
  y_start <- min(-30, sb$tumor_day-30, na.rm=TRUE)
  par(mfrow=c(1,2))
  par(mar=c(2.6, 2.6, 2.6, 0.6))
  plot(sb$cumulative_probability, sb$days, type="S", xlim=c(0,1), ylim=c(sb$days[n]+30, y_start))
  # Note: if there is no tumor_day (as for some of the controls, for example),
  # then the following two lines of code will quietly fail, and produce no arrow
  # which is exactly what we want.
  arrows(0.5, sb$tumor_day[1], 0.1, sb$tumor_day, lwd=3, length=0.1, col="red")
  text(0.55, sb$tumor_day[1], "Registry day", col="red", adj=0)
  segments(0.1, -30, 0.1, 0)
  segments(0.1, 0, sb$cumulative_probability[1], 0)
  segments(sb$cumulative_probability[n], sb$days[n], sb$cumulative_probability[n], sb$days[n]+30)
  title(paste(sb$patient_num[1], ", ", sb$GP[1], ", ", res, sep=""))
}
write_time_course <- function(sb) {
  n <- dim(sb)[1]
  y_start <- min(-30, sb$tumor_day-30, na.rm=TRUE)
  par(mar=c(2.6, 0, 2.6, 0))
  plot(sb$cumulative_probability, sb$days, type="n", xlim=c(0,1), ylim=c(sb$days[n]+30, y_start), axes=FALSE)
  title(paste("First event on ", sb$observation_date[1], sep=""))
  text(0, sb$days, sb$dx_label, cex=1, adj=0)
  cat(paste(c(sb$patient_num[1], sb$GP[1], res), collapse=", "))
  cat("\n")
  print(sb[, c("days", "co", "cumulative_probability", "dx_label")])
}


true_positive_patients <-(md$probability > 0.90) & (md$GP==i_case)
res <- "True Pos."
for (p in sample(md$patient_num[true_positive_patients], 20)) {
  sb <- calculate_time_course(p)
  plot_time_course(sb)
  write_time_course(sb)
}  

false_positive_patients <- (md$probability > 0.90) & (md$GP!=i_case)
res <- "False Pos."
for (p in sample(md$patient_num[false_positive_patients], 20)) {
  sb <- calculate_time_course(p)
  plot_time_course(sb)
  write_time_course(sb)
}  

false_negative_patients <- (md$co > 0) & (md$probability < 0.90) & (md$GP==i_case)
res <- "False Neg."
for (p in sample(md$patient_num[false_negative_patients], 20)) {
  sb <- calculate_time_course(p)
  plot_time_course(sb)
  write_time_course(sb)
}  

true_negative_patients <- (md$co > 0) & (md$probability < 0.90) & (md$GP!=i_case)
res <- "True Neg."
for (p in sample(md$patient_num[true_negative_patients], 20)) {
  sb <- calculate_time_course(p)
  plot_time_course(sb)
  write_time_course(sb)
}  
```

Save the important stuff

```{r save_important, timer=TRUE}
save(mc, md, me, file="evaluate_time_course.RData")
```

Save everything

```{r save_everything, timer=TRUE}
save.image(file="evaluate_model.RData")
```

Well done. Here is how long everything took.

```{r display_timing_log}
if (qc) {
  cat("Program began at ")
  cat(as.character(start_time))
  cat("\nProgram ended at ")
  cat(as.character(Sys.time()))
  cat("\n\n")
  tm <- read.table("timing_log.txt")$V1
  cat(paste(tm, collapse="\n"))
}
```
