```{r create_basic_documentation, echo=TRUE}
doc <- NULL
doc$run.date <- date()
doc$version <- system(' git rev-parse HEAD', intern=TRUE)
doc$author <- "Steve Simon (KUMC)"
doc$maintainer <- "Steve Simon (KUMC)"
doc$assistants <- "Dan Connolly"
```


Case-control
============
For context, see [485].

[485]: https://informatics.gpcnetwork.org/trac/Project/ticket/485
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

This program reads data from PCRONET CDM and matches it
with patient numbers from i2b2 to get a case-control
data set. It stores the data and some intermediate files in
an RData file.

Here's a summary of how the data was selected. We
identified four disease groups in Heron (KUMC's i2b2).

Breast cancer (as identified in the SEER Site Summary),

All other cancers (again from SEER),

Diabetes (ICD9 250 or ICD10 E08-E13), and

Ischemic heart disease (ICD9 410-414.99 or ICD10 I20-I25).

All males were excluded.

The first group represents cases and the other three
represent different control groups. Patients who qualified
for two or more of the above groups were excluded from the
analysis.

There were no date restrictions. In future versions, we may want
to consider some date restrictions and try matching subjects on
age.

We then merged that data with the PCORNET CDM to get 
any ICD9 codes associated with each patient.

Two companion programs, analyze_lasso.Rmd and analyze_sensitivity.Rmd,
take the resulting data set and produces simple graphs and analyses. 

This program was run on `r doc$date` using version `r doc$version`.
The original author is `r doc$author`. `r doc$maintainer`
is currently maintaining and enhancing this program
with the assistance of `r doc$assistants`.


```{r load_required_libraries, echo=TRUE}
library("ROracle")
library("RSQLite")
save.image("backup.RData")
rm(list=ls())
verbose=TRUE # if TRUE, print various intermediate values and quality checks.
if (verbose) {
  cat("\n\nQuality check: Are we in the correct directory?")
  print(getwd())
}
# Don't wrap so much
options(width=90)
start_time <- Sys.time()
if (verbose) {
  cat("\n\nProgram started at ")
  print(start_time)
}
display_elapsed_time <- function() {
  cat("\n\nThis code chunk required ")
  print(round(Sys.time()-start_time))
  return(Sys.time())
}
if (verbose) {
  cat("\n\nQuality check: Are we in the correct directory?")
  print(getwd())
}
# Don't wrap so much
options(width=90)
```

Here are the special functions needed for this program and brief
tests of their functionality.

```{r create_special_functions}
align_icd9_codes <- function(dx_old) {
  # This program strips out the dots and pads short
  # icd9 codes with zeros. It prints out anything
  # that is too short or too long.
  dx_new <- gsub(".","",dx_old,fixed=TRUE)
  dx_new <- ifelse(nchar(dx_new)==3,paste(dx_new,"0",sep=""),dx_new)
  dx_new <- ifelse(nchar(dx_new)==4,paste(dx_new,"0",sep=""),dx_new)
  cat("\n\nThe following diagnosis codes are invalid:\n")
  print(sort(unique(dx_new[nchar(dx_new)!=5])))
  cat(".\n")
  return(dx_new)
}
if (verbose) {
  cat("\nSimple test.\n")
  align_icd9_codes(c("0014","V28.9","XX"))
}
list_random_rows <- function(df,n=5) {
  # select and return first five rows, last five rows, and
  # a few random rows from a data frame.
  # If input is a vector, coerce it into a data frame
  df <- data.frame(df, stringsAsFactors=FALSE)
  new_list <- NULL
  nrows <- dim(df)[1]
  if (nrows<=3*n) {return(list(All_rows=df))}
  top_name <- paste("First",n,"rows",sep="_")
  mid_name <- paste("Random",n,"rows",sep="_")
  bot_name <- paste("Last",n,"rows",sep="_")
  new_list[[top_name]] <- head(df,n)
  selected_rows <- sample((n+1):(nrows-n),n)
  new_list[[mid_name]] <- df[selected_rows, ]
  new_list[[bot_name]] <- tail(df,n)
  return(new_list)
}
print_random_rows <- function(df, n=5) {
  print(list_random_rows(df, n))
}
if (verbose) {
  cat("\nSimple test.\n")
  print_random_rows(1:100)
  print_random_rows(c(LETTERS,letters))
}
strip_specials <- function(x0) {
  # This function strips special characters from a character vector,
  # replacing most of them with an underscore.
  x0 <- gsub(" ","_",x0,fixed=TRUE)
  x0 <- gsub("-","_",x0,fixed=TRUE)
  x0 <- gsub('"',"_", x0,fixed=TRUE)
  x0 <- gsub('/',"_", x0,fixed=TRUE)
  x0 <- gsub('+',"_", x0,fixed=TRUE)
  x0 <- gsub('&',"_", x0,fixed=TRUE)
  return(x0)
}
if (verbose) {
  cat("\nSimple test.\n")
  strip_specials(c("test one","test-two","test&three"))
  start_time <- display_elapsed_time()
}
```

Set up storage for temporary data frames and matrices (d)
and notes for my own reference.

```{r setup_temp}
d <- NULL
notes <- NULL
if (verbose) start_time <- display_elapsed_time()
```

This program also relies on data found in the PCORnet 
Common Data Model. Here is the code needed to access
that information and a simple test.

```{r read_config, echo=TRUE, error=FALSE}
# This code borrowed from cdm_fun.Rmd.
cdm_config <- read.csv('../cdm_config.csv', stringsAsFactors=FALSE)
missing_config <- setdiff(c('account', 'password'), names(cdm_config))
stopifnot(length(missing_config) == 0)
cdm <- dbConnect(Oracle(), cdm_config$account, cdm_config$password, cdm_config$access)
if (verbose) {
  cat("Simple test")
  dbGetQuery(cdm, "select * from pcornet_cdm.diagnosis where rownum < 10")
  start_time <- display_elapsed_time()
}
```


This section reads in the i2b2 files.

The file, use-this-database.txt, should have one entry for each file.
It also has information about which site the data 
comes from and which disease group the patients come from.

```{r setup_sql_queries, echo=TRUE}
setwd("/d1/home/ssimon/bc_qa/bc_case_control")
d$fi <-
  read.csv(file="use_this_database.txt", header=TRUE, stringsAsFactors=FALSE)
notes$fi <- "File information"
if (verbose) print(d$fi)
i.case <- d$fi$disease_group[d$fi$case==1]
i.control <- d$fi$disease_group[d$fi$case==0]
if (verbose) start_time <- display_elapsed_time()
```

Now design the appropriate sql queries.

```{r design-sql-queries}
sql <- NULL

# get distinct patient numbers in each disease group.
sql[["distinct_patients"]] <-
  "select distinct patient_num
   from observation_fact"

sql[["matching_count"]] <- 
  "select distinct j.PATIENT_NUM, j.GP, p.PATID
   from pcornet_cdm.diagnosis P
   JOIN DP j
   on p.PATID = j.PATIENT_NUM
   where p.DX_TYPE = '09'"

sql[["patient_diagnoses"]] <-
  "select distinct j.PATIENT_NUM, j.GP, p.DX
   from pcornet_cdm.diagnosis P
   join DP j
   on p.PATID = j.PATIENT_NUM
   where p.DX_TYPE='09'"

if (verbose) start_time <- display_elapsed_time()
```

Apply these queries to each group. The dimensions in dp
should match the counts in nd.

```{r get_overall_patient_counts}
# dp: distinct patient numbers in each disease group
dp <- NULL
file_names <- d$fi$file_location
group_names <- d$fi$disease_group
if (any(group_names!=sort(group_names))) {
  cat("\\n\nWarning: Data should be sorted by disease group")
}
for (i in 1:length(file_names)) {
  file_name <- file_names[i]
  group_name <- group_names[i]
  if (verbose) {
    cat("\nRunning queries for ", group_name, " (", file_name,  ").\n", sep="")
  }
  conn_site <- dbConnect(SQLite(), dbname=file_name)
  dp[[group_name]] <- dbGetQuery(conn_site, sql["distinct_patients"])
}
if (verbose) {
  sapply(dp,dim)
  start_time <- display_elapsed_time()
}
```

Stack all the files and note which is in which group.

```{r stack_patient_numbers}
d$st <- data.frame(gp=names(dp[1]), dp[[1]], stringsAsFactors=FALSE)
notes$st <- "Stack the individual patient numbers in a single data frame"
for (k in 2:length(dp)) {
  d$st <- rbind(d$st, data.frame(gp=names(dp[k]), dp[[k]]))
}
if (verbose) {
  print_random_rows(d$st)
  start_time <- display_elapsed_time()
}
```

Next, you need to write the patient numbers to the same
location as the PCORnet CDM database.

```{r write_distinct_patient_numbers}
names(d$st) <- toupper(names(d$st))
dbWriteTable(cdm, "DP", d$st, overwrite=TRUE)
if (verbose) {
  dbListTables(cdm)
  start_time <- display_elapsed_time()
}
```

Now pull out all the diagnosis codes associated with our patient list.

Note: there are patients who are in the various
disease groups, but who do not have any ICD9 codes.
I need to investigate why.

```{r pull_out_diagnoses}
mc <- dbGetQuery(cdm, sql[["matching_count"]])
n1 <- table(mc$GP) 
n2 <- table(d$st$GP)
if (verbose) print(c(n1,n2))
d$di <- dbGetQuery(cdm, sql[["patient_diagnoses"]])
notes$di <- "SQL query from pcornet.cdm"
if (verbose) {
  print(dim(d$di))
  print_random_rows(d$di)
  start_time <- display_elapsed_time()
}
```

Now, let's allow ourselves the ability to screen out
any diagnosis codes that occur infrequently among the
cases.

```{r find_common_dx}
# Add a dx_count column that notes how often the DX appears
# among the case groups. This will allow you to select only
# the "popular" DX codes.
d$tb <- table(d$di$DX[d$di$GP==i.case])
notes$tb <- "Frequency counts for DX codes among the cases."
# Note: as.numeric is needed here because you need to strip out
# the names before merging.
# Also note that this merge statement will remove any diagnosis
# codes which occur only among the controls.
d$co <- data.frame(DX=names(d$tb), dx_count=as.numeric(d$tb))
notes$co <- "Convert table into data frame"
d$cd <- merge(d$di, d$co)
notes$cd <- "Merge frequency counts in with original diagnoses."
if (verbose) {
  print_random_rows(d$co)
  print_random_rows(d$cd)
  print(length(unique(d$di$DX)))
  print(length(unique(d$cd$DX)))
  start_time <- display_elapsed_time()
}
```

While you can get the nice labels for icd9 codes from i2b2, it is easier
just to pull them from another source. I chose labels from the
[https://www.cms.gov/medicare/coding/ICD9providerdiagnosticcodes/codes.html CMS]
site. 

```{r get-nice-names, echo=TRUE}
# il = icd9_labels
icd9_file <- "icd9_labels.csv"
d$il <- read.csv(icd9_file, header=TRUE, as.is=TRUE, row.names=NULL)
notes$il <- "icd9 codes and labels"
# Double check to make sure nothing was converted to factors.
d$il$nice_label <- paste(strip_specials(d$il$short_label),d$il$DX,sep="_")
if (verbose) print_random_rows(d$il$nice_label)
# Here are diagnosis codes in cd, not found in il.
if (verbose) print_random_rows(sort(setdiff(d$cd$DX,d$il$DX)))
d$il$dx_new <- align_icd9_codes(d$il$diagnosis_code)
d$cd$dx_new <- align_icd9_codes(d$cd$DX)
# Here are diagnosis codes in cd, not found in il.
if (verbose) print(sort(setdiff(d$cd$dx_new,d$il$dx_new)))
d$il$dx_label <- paste(strip_specials(d$il$short_label), d$il$dx_new, sep="_")
d$lb <- merge(d$cd, d$il[, c("dx_new","dx_label")], all.x=TRUE, all.y=FALSE)
notes$lb <- "DX codes and nice labels"
if (verbose) {
  print_random_rows(d$lb)
  start_time <- display_elapsed_time()
}
```

Now save everything for later use.

```{r save-everything, echo=TRUE}
save.image(file="extract_case_control.RData")
lb <- d$lb
st <- d$st
save(
  lb, st, i.case, i.control,
  list_random_rows, print_random_rows
  file="case_control_data.RData")
if (verbose) {
  start_time <- display_elapsed_time()
  cat("\n\nProgram ended at")
  print(Sys.time())

```
