# Breast Cancer All-Site Datamart - Consented Patients Only
```{r echo=FALSE}
# ============================================================================
###  Creates site-specific tumor-level datamart
# Input Dataframes
#   dataset       - desriptors for site's input dataset 
#   tumor.site    - pt tumors listing
#   survey.sample - complex frame created by bc_excl
# 
# Note - script was modeled after 'bc_random_smaple.rmd'
#
# 25-Nov Genesis (Parceled out f/ original bc_site_datamart)
# 30-Nov SAVEPOINT (RMD file committed to TortoiseHg)
# ============================================================================
```
#### Summary Statistics
#### Data Dictionary
#### Table 1 Statistics (All Variables)
#### Descriptive Analysis (For Each Variable)
#### Cross-Site Analysis of Key Variables

### Study Exclusion Criteria
#### - Sex not equal to female
#### - Less than 18 years of age
#### - Prior cancer diagnosis
#### - Breast cancer was not microscopically confirmed
#### - Only tumor morphology was lobular carcinoma in situ
#### - Stage IV breast cancer
#### - Known to be deceased

```{r Package Set-Up, include=FALSE}

# Include function libraries
source('/d1/home/vleonardo/PHM-Development/PHM-LIBRARY.rmd')  # Loads PHM functions
source('/d1/home/vleonardo/GPC-Development/bc-datamart/bc_dm_library.R')  # Loads PHM functions

PHM.PackageSetup()
library(RColorBrewer)

# Don't wrap so much
options(width=300)
opts_chunk$set(echo=FALSE)  # PDF creates better sized charts 

```

### Summary Statistics

``` {r Summary Statistics}
v.consented.datamart <- read.csv("/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-DatamartOutput/BCDatamart-ALLSITE-CONSENTED.csv",stringsAsFactors=FALSE)
BCRup.Summary.Statistics(v.consented.datamart)
```

### Data Dictionary

```{r Data Dictionary, echo=FALSE, eval=TRUE}

message("RECORD IDENTIFIERS & ELIGIBILTY CRITERIA")
print(str(v.consented.datamart[1:25]))

message("NAACCR VARIABLES")
v.col.terms <- BCRup.Initialize.Col.Terms()
print(str(v.consented.datamart[26:ncol(v.consented.datamart)]))
```

### Table 1 Statistics  (All Variables)

```{r Datamart Table 1 - Consented, echo=FALSE, eval=TRUE}
# Provides a consolitedate presentation similar to a standard 'Table One' 
for (tmp.col.index in 7:(ncol(v.consented.datamart))) # Skip P#, E# and Site columns
  {  
  tmp.col.name <- names(v.consented.datamart[tmp.col.index])
  BCRup.DescribeDatamartColumn(v.consented.datamart, tmp.col.name)  
  }
```

### Descriptive Analysis (For Each Variable)

```{r Datamart Descriptive Analysis - Consented, echo=FALSE, eval=TRUE}

for (tmp.col.index in 7:(ncol(v.consented.datamart))) # Skip P#, E# and Site columns
  {  
  tmp.col.name <- names(v.consented.datamart[tmp.col.index])                         
  #BCRup.DescribeDatamartColumn(v.consented.datamart, tmp.col.name) 
  BCRup.VisualizeDatamartColumn(v.consented.datamart,tmp.col.name)
  BCRup.DescribeDatamartColumn(v.consented.datamart, tmp.col.name)  
  }
```

### Cross-Site Analysis of Key Variables

```{r Visual Analysis - Consented, echo=FALSE, eval=TRUE}

#################################################################
# Set-up for site-breakdowns
v.site.cnt <- length(unique(v.consented.datamart$gpc.site.name))

# Establish color palette
v.gpc.site.colors <- c(blues9[9],"#A40047",'#008A60',"#EFB000","#790018","#AC1F2D","#013972","#A7003B")
v.vpl.blues <- brewer.pal(9, "Blues")

#################################################################
# Consented Patients By Site
p <- unique(v.consented.datamart[,c("gpc.site.name","patient.num")])

ggplot(p, aes(x=gpc.site.name,ymax=..count..))  + 
  ggtitle("Consented Patients") +
  labs(x="",y="") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=32, hjust=0),
        axis.text.x = element_text(size=18,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
#  geom_bar(width=.5,  fill=blues9[(10-v.site.cnt):9]) +  # Generates n rgb values, plus an offset towards darker side of the scale
  geom_bar(width=.5,  fill=v.gpc.site.colors) +  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150)) +  # Extend y-axis so values don't get cropped
  stat_bin(geom="text",aes(label=..count..),vjust=-0.5) # Display values above bars

#################################################################
# Encounters By Site
p <- v.consented.datamart
ggplot(p, aes(x=gpc.site.name,ymax=..count..))  + 
  ggtitle("Encounters (Tumors)") +
  labs(x="",y="") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=32, hjust=0),
        axis.text.x = element_text(size=18,face="bold",color="darkslategray")) +
  geom_bar(width=.5,  fill=v.gpc.site.colors) +  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150)) +  # Extend y-axis so values don't get cropped
  stat_bin(geom="text",aes(label=..count..),vjust=-0.5) # Display values above bars

#################################################################
# ER1
tmp.col.name <- "NAACCR.2880.ER.CSSSF01"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]
colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Coded As Missing"
p$tmp.col.descriptor[is.na(p$tmp.col.name)] <- "Code Not Provided"
p[grepl("10",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Positive"
p[grepl("20",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Negative"
p[grepl("30",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Borderline"
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Positive","Negative","Coded As Missing","Code Not Provided"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c("darkred","darkgreen","darkgray","aliceblue")) +
  ggtitle("CSSSF01 - ER (Estrogen Receptor)") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +  
  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

#################################################################
# PR1
tmp.col.name <- "NAACCR.2890.PR.CSSSF02"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]
colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Coded As Missing"
p$tmp.col.descriptor[is.na(p$tmp.col.name)] <- "Code Not Provided"
p[grepl("10",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Positive"
p[grepl("20",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Negative"
p[grepl("30",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Borderline"
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Positive","Negative","Coded As Missing","Code Not Provided"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c("darkred","darkgreen","darkgray","aliceblue")) +
  ggtitle("CSSSF02 - PR (Progesterone Receptor)") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +  
  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

##############################################################
# HER2
tmp.col.name <- "NAACCR.2869.HER2.CSSSF15"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]
colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Coded As Missing"
p$tmp.col.descriptor[is.na(p$tmp.col.name)] <- "Code Not Provided"
p[grepl("10",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Positive"
p[grepl("20",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Negative"
p[grepl("30",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Borderline"
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Positive","Borderline","Negative","Coded As Missing","Code Not Provided"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c("darkred","red","darkgreen","darkgray","aliceblue")) +
  ggtitle("CSSSF15 - HER2 (Human Epidermal Growth Factor Receptor 2)") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

##############################################################
# Muligene Signature Method
tmp.col.name <- "NAACCR.2876.MS.Method.CSSSF22"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]

colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Coded As Missing"
p$tmp.col.descriptor[is.na(p$tmp.col.name)] <- "Code Not Provided"
p[grepl("10",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Oncotype DX"
p[grepl("20",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "MammaPrint"
p[grepl("30",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Other"
p[grepl("40",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Other"
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Oncotype DX","MammaPrint","Other","Coded As Missing","Code Not Provided"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c(blues9[9],"lightsteelblue",blues9[8],"darkgray","aliceblue")) +
  ggtitle("CSSSF22 - Multigene Signature Method") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  # Generates n rgb values, plus an offset towards darker side of the scale
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

##############################################################
# MGS Result
tmp.col.name <- "NAACCR.2877.MS.Result.CSSSF23"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]

colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Actl Score Provided"
p$tmp.col.descriptor[is.na(p$tmp.col.name)] <- "Score Not Provided"
p[grepl("200",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Low Risk"
p[grepl("300",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Intermediate Risk"
p[grepl("400",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "High Risk"
p[grepl("988|997|998|999",p[,"tmp.col.name"]),"tmp.col.descriptor"] <- "Coded As Missing"
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Actl Score Provided","Low Risk","Intermediate Risk","High Risk","Coded As Missing","Score Not Provided"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c(blues9[9],"darkgreen","red","darkred","darkgray","aliceblue")) +
  ggtitle("CSSSF23 - Multigene Signature Score") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

##############################################################
# 380 Seqno Central
tmp.col.name <- "NAACCR.0380.Seqno.Central.Descriptor"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]

colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- p$tmp.col.name

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
#  scale_fill_manual(values=c(blues9[9],"darkgreen","red","darkred","darkgray","aliceblue")) +
  ggtitle("0380 - Sequence Number Central Descriptor") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped

##############################################################
# 560 Seqno Hospital
tmp.col.name <- "NAACCR.0560.Seqno.Hosp.Descriptor"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name)]

colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- p$tmp.col.name

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
#  scale_fill_manual(values=c(blues9[9],"darkgreen","red","darkred","darkgray","aliceblue")) +
  ggtitle("0560 - Sequence Number Hospital") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped


##############################################################
# Sequence Number using 'aligned' i2b2 terms
tmp.col.name1 <- "NAACCR.0380.Seqno.Central"
tmp.col.name2 <- "NAACCR.0560.Seqno.Hosp"
p <- v.consented.datamart[,c("gpc.site.name",tmp.col.name1,tmp.col.name2)]

#colnames(p)[colnames(p)==tmp.col.name] <- "tmp.col.name"
p$tmp.col.descriptor <- "Code Not Recognized"
p$tmp.col.descriptor[grepl("00",p[,tmp.col.name1]) | grepl("00",p[,tmp.col.name2])] <- "Only 1 Primary"
p$tmp.col.descriptor[grepl("01",p[,tmp.col.name1]) | grepl("01",p[,tmp.col.name2])] <- "1st of Mult Primaries"
p$tmp.col.descriptor[grepl("02|03",p[,tmp.col.name1]) | grepl("02|03",p[,tmp.col.name2])] <- "Other Mult Primaries"
p$tmp.col.descriptor[grepl("60",p[,tmp.col.name1]) | grepl("60",p[,tmp.col.name1])] <- "Non-Malignant"

# Order stacked bar values
p$tmp.col.descriptor <- factor(p$tmp.col.descriptor,levels=c("Only 1 Primary","1st of Mult Primaries","Other Mult Primaries","Non-Malignant","Code Not Provided","Coded Not Recognized"))

ggplot(p, aes(x=gpc.site.name,fill=tmp.col.descriptor))  + 
  geom_bar(stat="bin",width=.75,color="darkgray") +  # 'bin' counts cases, 'identity' represents values
  scale_fill_manual(values=c("darkred","red","indianred","darkgreen","darkgray","yellow")) +
  ggtitle("'Aligned' Sequence Number (0380 Central & 0560 Hospital)") +
  labs(x="",y="Encounters (Tumors)") +  # Labels for x & y axes
  theme(plot.title = element_text(color="darkslategray", face="bold", size=24, hjust=0),
        axis.text.x = element_text(size=12,face="bold",color="darkslategray"),
        axis.text.y = element_text(size=12,face="bold"),
        legend.title=element_blank()) +
  guides(fill = guide_legend(reverse=TRUE)) +
  scale_y_continuous(limits = c(0,150))  # Extend y-axis so values don't get cropped
```

