# Breast Cancer - All-Site Datamart
###    - Filters dataset to only include consented patients
```{r echo=FALSE}
# ============================================================================
# 30-Nov Genesis (Parceled out f/ original bc_site_datamart)
# 24-Nov SAVEPOINT (RMD file committed to TortoiseHg)
# ============================================================================
```

### Package Set-Up and Initialization 

```{r Package Set-Up, include=FALSE}

# Include PHM function libraries
source('/d1/home/vleonardo/PHM-Development/PHM-LIBRARY.rmd')  # Loads PHM functions
# library(ggplot2)
library(reshape)
PHM.PackageSetup()

# Don't wrap so much
options(width=300)
opts_chunk$set(echo=FALSE)

# BC Function library
source('/d1/home/vleonardo/GPC-Development/bc-datamart/bc_dm_library.R')  
```

### STAGE 2 - BUILD SITE-SPECIFIC CONSENTED DATAMARTS (ONE AT A TIME)

```{r Build site-specific consented datamarts -- One at a time}

v.output.dir <- '/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-DatamartOutput/'  
v.sites <- c('WISC')  # Originally, was a list of all sites
message("Getting consented patients for ",v.sites[1])
v.site.consented.datamart <- BCRup.GetConsentedPtData(v.sites[1])

# Final packaging for output
v.site.consented.datamart$gpc.sex_cd <- NULL
v.site.consented.datamart$gpc.birth_date <- NULL
v.site.consented.datamart <- v.site.consented.datamart[order(
  v.site.consented.datamart$gpc.study.id,
                                                                     v.site.consented.datamart$encounter.num),]
v.filename <- paste0(v.output.dir,"BCDatamart-Site-Consented-",v.sites[1],".csv")
write.csv(v.site.consented.datamart, file = v.filename, row.names=FALSE)
message('Number of pts consented: ',length(unique(v.site.consented.datamart$patient.num)))
message('Number of rows written: ',nrow(v.site.consented.datamart))

```

### STAGE 3 - ALL-SITE DATAMART, CONSOLIDATE 

```{r Consolidate and filter for consented study ids}

v.output.dir <- '/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-DatamartOutput/'  
#v.sites <- c('KUMC','MCRF','MCW','UIOWA','UMN','UNMC','UTSW','WISC')
v.sites <- c('KUMC','MCRF','UIOWA','UMN','UTSW','WISC')
v.filename <- paste0(v.output.dir,"BCDatamart-Site-Consented-",v.sites[1],".csv")
v.consented.datamart <- read.csv(v.filename,stringsAsFactors=FALSE)
message("... Rows added for ",v.sites[1])
for (i in v.sites[2:length(v.sites)]) {
  v.filename <- paste0(v.output.dir,"BCDatamart-Site-Consented-",i,".csv")
  tmp.site.consented.datamart <- read.csv(v.filename) 
  v.consented.datamart <- rbind(v.consented.datamart,tmp.site.consented.datamart)
  message("... Rows added for ",i)
}
message(' ')
print(setNames(aggregate(cbind(patient.num,encounter.num)~gpc.site.name,
                data=v.consented.datamart,
                function(x) length(unique(x))),
          c("GPC Site","Patients","Enctr(Tumors)")))

# Export 'All-Site Consented' datamart
v.consented.datamart <- v.consented.datamart[
  order(v.consented.datamart$gpc.site.name,
        v.consented.datamart$gpc.study.id,
        v.consented.datamart$encounter.num),]
v.filename <- paste0(v.output.dir,"BCDatamart-ALLSITE-CONSENTED.csv")
write.csv(v.consented.datamart, file = v.filename, row.names=FALSE)
message(' ')
message('Total rows written: ',
        nrow(v.consented.datamart), " enctrs/tumors  ",
        length(unique(v.consented.datamart$patient.num))," pts")

```

### STAGE 4 - APPLY STANDARDIZED DECODING FOR SELECTED COLUMNS (Columns where sites did not provide code names)
# Note - Decoding is no longer performed
```{r Decoding for select columns}
# v.codebook <- read.csv("/d1/home/vleonardo/GPC-Development/bc-datamart/i2b2 codebook-decoded-subset.csv",stringsAsFactors=FALSE)
# v.naaccr.code.list <- c('0610','0670','0820','0830','2850','2860','2869','2876','2877','2880','2890','2940','3000','3020','3400','3430')
# v.consented.datamart <- read.csv("/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-DatamartOutput/BCDatamart-ALLSITE-CONSENTED.csv",stringsAsFactors=FALSE)
# for (i.naaccr in v.naaccr.code.list) { 
#   tmp.col.name <- as.character(v.col.terms$col.name[grepl(i.naaccr,v.col.terms$col.name)])
#   message('... decoding: ',i.naaccr,'-',tmp.col.name)
#   tmp.col.descriptor <- paste0(v.col.terms$col.name[grepl(i.naaccr,v.col.terms$col.name)],'.Descriptor')
#   v.consented.datamart$code_value <- as.character(v.consented.datamart[,tmp.col.name])
# #  v.consented.datamart$label <- unique(v.codebook$label[v.codebook$code_value==v.consented.datamart$code_value])
#   v.consented.datamart <- merge(v.consented.datamart,v.codebook[,c("code_value","label")],
#         by.x=c("code_value"), by.y=c("code_value"),all.x=TRUE)
#   v.consented.datamart[,tmp.col.descriptor] <- v.consented.datamart$label
#   v.consented.datamart$code_value <- NULL
#   v.consented.datamart$label <- NULL
# 
# }
# 
# # Export 'All-Site Consented' datamart
# v.consented.datamart <- v.consented.datamart[with(v.consented.datamart,order(gpc.site.name,gpc.study.id, encounter.num)),]
# v.filename <- paste0(v.output.dir,"BCDatamart-ALLSITE-CONSENTED-DECODED.csv")
# write.csv(v.consented.datamart, file = v.filename, row.names=FALSE)
# message('Number of rows written: ',nrow(v.consented.datamart))

```

### ALL-SITE DATAMART - Summary Statistics

``` {r Summary Statistics}
BCRup.Summary.Statistics(v.consented.datamart)
```
