# Breast Cancer - Site Datamart
```{r echo=FALSE}
# ============================================================================
# Creates site-specific tumor-level datamart (staging data)
# Input Dataframes
#   dataset       - desriptors for site's input dataset 
#   tumor.site    - pt tumors listing
#   survey.sample - complex frame created by bc_excl
# 
# 30-Nov Genesis (Parceled out f/ original bc_site_datamart)
# 30-Nov SAVEPOINT (RMD file committed to TortoiseHg)
#
# ============================================================================
```

### Package Set-Up and Initialization 

```{r Package Set-Up, include=FALSE}

# Include PHM function libraries
source('/d1/home/vleonardo/PHM-Development/PHM-LIBRARY.rmd')  # Loads PHM functions

# library(ggplot2)
library(reshape)
library(RColorBrewer)
PHM.PackageSetup()

# Don't wrap so much
options(width=300)
opts_chunk$set(echo=FALSE)

# BC Function libraries
source('/d1/home/vleonardo/GPC-Development/bc_qa/bc_qa_txform.R')  
source('/d1/home/vleonardo/GPC-Development/bc-datamart/bc_dm_library.R')  

# Load global dataframes
v.site.files <<- BCRup.Initialize.Site.Files()
v.col.terms  <<- BCRup.Initialize.Col.Terms()
```


### Loading 'dataset' object

```{r}
#input <- source('dataset.R')$value

# conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, submitter $name, $issues_other
# dataset <- input$about

v.output.dir <- '/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-DatamartOutput/'
v.site <- "UTSW"
dataset <- list("")
dataset$site <- v.site
#dataset$bc_db <- "KUMC-16-kumcBC"
#dataset$bc_db <- "UMN-30-UMN BC SQLLite v2 with seq num"
dataset$bc_db <- v.site.files[v.site.files$site==v.site,"source.db"]

message('Dataset identified: ',dataset$bc_db)
conn.site <- dbConnect(SQLite(),paste0('/d1/home/vleonardo/GPC-Development/bc-data-files/Phase1-SourceDbs/',dataset$bc_db) )
print(dbListTables(conn.site))

# builder.summary reads raw db
# Note - this is not used for the concept terms, rather 'code paths' from the concept dimension is what is used.
bs <- (builder.summary(conn.site)[, c('variable', 'pat_qty', 'enc_qty', 'fact_qty')])
print(bs)

# Load site's variable and concept codes directly from db file
site <- list("")
site$variables <- dbGetQuery(conn.site,"select * from variable")
print(site$variables[,c("concept_path","name_char")])
site$concept.codes <- dbGetQuery(conn.site,"select * from concept_dimension")

# Diagnostics for site's data
message('SQLITE db file) patient_dimension: ',
        dbGetQuery(conn.site,"select count(*) from patient_dimension"))
message('SQLITE db file) variable: ',
        dbGetQuery(conn.site,"select count(*) from variable"))
message('SQLITE db file) concept_dimension: ',
        dbGetQuery(conn.site,"select count(*) from concept_dimension"))
message('SQLITE db file) observation_fact: ',
        dbGetQuery(conn.site,"select count(*) from observation_fact"))

```

### Load and Merge Observations With Concept Descriptors

``` {r}
############################################################################
# Delaware: n=4660583
# Expected Run-time: 6.5 mins
site$ptobs <- PHM.LoadObservationsWithConceptDescriptors(conn.site)

#saveRDS(site$ptobs,  paste0(v.output.dir,"BCDatamart-PtObs-",dataset$site,".rds"))
print(nrow(site$ptobs))
site$ptobs.code.paths <- unique(site$ptobs$code.path)                           
```

### Loading original patient data for `r dataset$site`
 -- variables used in QA process (exclusion criteria and receptor status)

```{r}
load("/d1/home/vleonardo/GPC-Development/bc_qa/bc_terms_results.RData")
message('Number of terms loaded: ',nrow(bcterm$term204))

tumor.site <- bc.exclusions(conn.site)   

print(names(tumor.site))
saveRDS(tumor.site,  paste0(v.output.dir,dataset$bc_db,"-",Sys.Date(),"-tumor-site.rds"))
x.tumor.site <- tumor.site  # Holding area for restoration/debugging purposes
#tumor.site <- readRDS(paste0(dataset$bc_db,"-",Sys.Date(),"-tumor-site.rds"))
message('Number of tumors loaded: ',(nrow(tumor.site)))
message('Number of pts in dataset: ',length(unique(tumor.site$patient_num)))


survey.sample <- check.cases(tumor.site)  # Creates tumor-level criteria flags
survey.sample$all.criteria <- reduce.logical(survey.sample)
message('Number of pts meeting inclusion criteria: ', 
        nrow(survey.sample[survey.sample$all.criteria,]))

survey.sample.summary <- count.cases(survey.sample)  # Produce table of criteria counts
print(survey.sample.summary)
```

### Fetch birth_date and sex for any patients w/o value

```{r}
# Note - tumor.site can have NAs for some birth_dates (possibly ineligible encounters)
if (nrow(subset(tumor.site,is.na(date.birth))) > 0) {
  nrow(subset(tumor.site,is.na(date.birth)))
  tmp.patient <- dbGetQuery(conn.site,"select patient_num, birth_date, sex_cd from patient_dimension")
  tumor.site <- merge(tumor.site,tmp.patient)
  tumor.site$date.birth <- tumor.site$birth_date
  tumor.site$sex <- tumor.site$sex_cd
  nrow(subset(tumor.site,is.na(date.birth)))
}
```

### Filter for only eligible encounters(tumors)
```{r}
# Remove original receptor status variables f/ inclusion set, as they are replicated in NAACCR vars
q.site.datamart <- subset(tumor.site,select=-c(er.csf.1,pr.csf.2,her2.csf.15,mgs.method.csf.22,mgs.score.csf.23))
colnames(q.site.datamart)[colnames(q.site.datamart)=="patient_num"] <- "patient.num"
colnames(q.site.datamart)[colnames(q.site.datamart)=="encounter_num"] <- "encounter.num"

# Add prefix to survey variables to distinguish them from TR & EHR direct-sourced items
for (i in 3:ncol(q.site.datamart)) {
  tmp.col.name <- names(q.site.datamart[i])
  colnames(q.site.datamart)[colnames(q.site.datamart)==tmp.col.name] <- paste0("gpc.",tmp.col.name)
}

# Add GPC Site Identifier
q.site.datamart <- cbind(gpc.site.name=dataset$site,q.site.datamart)

# Work-around for UIOWA) GPC.date.birth is NA for some pts, but sqldb has value in patient dimension

# Calculate age at diagnosis
tmp.dx <- list("")
tmp.dx <- q.site.datamart[,c("encounter.num","gpc.date.dx","gpc.date.birth")]
tmp.dx$age <-   
  as.POSIXlt(tmp.dx$gpc.date.dx)$year - 
  as.POSIXlt(tmp.dx$gpc.date.birth)$year
tmp.dx$dx.mmdd <- (as.POSIXlt(tmp.dx$gpc.date.dx)$mon*100) +
                             as.POSIXlt(tmp.dx$gpc.date.dx)$mday 
tmp.dx$birth.mmdd <- (as.POSIXlt(tmp.dx$gpc.date.birth)$mon*100) +
                             as.POSIXlt(tmp.dx$gpc.date.birth)$mday 
tmp.dx$age[!(is.na(tmp.dx$gpc.date.birth)) & (tmp.dx$dx.mmdd < tmp.dx$birth.mmdd)] <- 
  tmp.dx$age[!(is.na(tmp.dx$gpc.date.birth)) & (tmp.dx$dx.mmdd < tmp.dx$birth.mmdd)] - 1
q.site.datamart$gpc.dx.age <- tmp.dx$age[tmp.dx$encounter.num==q.site.datamart$encounter.num]
# Move column towards beginning of list
q.site.datamart <- q.site.datamart[,c(1,2,3,ncol(q.site.datamart),5:ncol(q.site.datamart)-1)]

# Add eligiblity flag
q.site.datamart <- merge(q.site.datamart,survey.sample[,c("patient_num","encounter_num","all.criteria")],
                         by.x=c("patient.num","encounter.num"),
                         by.y=c("patient_num","encounter_num"))
colnames(q.site.datamart)[colnames(q.site.datamart)=="all.criteria"] <- "gpc.enctr.eligible"

message("Total number of patients submitted: ", length(unique(q.site.datamart$patient.num)))
message("Total number of encounters submitted: ", length(unique(q.site.datamart$encounter.num)))
message("Number of eligible patients: ", length(unique(q.site.datamart$patient.num[q.site.datamart$gpc.enctr.eligible])))
message("Number of eligible encounters(tumors): ", nrow(q.site.datamart[q.site.datamart$gpc.enctr.eligible,]))
```

### Trim master patient observations for only eligible encounters, solely for performance reasons

```{r}
message('Number of original observations: ',nrow(site$ptobs))
site$ptobs <- subset(site$ptobs, encounter.num %in% q.site.datamart$encounter.num)
message('Number of observations for eligible pts: ',nrow(site$ptobs))
```


### Add variables, specifying 2 search terms for 'variable name'

```{r}
# Variable Initialization
v.script.log <- data.frame(p.new.col.name=NA,p.code.string.1=NA,p.code.string.2=NA,log.msg=NA,log.action.taken=NA)
message("Rows in tumor.site: ", nrow(tumor.site))
tmp.tumor.site <- q.site.datamart

# Add each variable
for (r in 1:nrow(v.col.terms)) {
  tmp.col.name <- as.character(v.col.terms$col.name[r])
  tmp.tumor.site <- BCRup.AddVariableToDatamart(tmp.tumor.site, 
                                                site$variables,
                                                site$ptobs, 
                                                v.col.terms$term.1[r],
                                                v.col.terms$term.2[r],
                                                tmp.col.name, 
                                                v.col.terms$col.data.type[r]) 
}
q.site.datamart <- tmp.tumor.site
v.script.log <- v.script.log[! (is.na(v.script.log$p.new.col.name)),]
```

###  Summary of Variables Loaded and Actions Taken

```{r Variable Summary Report} 
names(q.site.datamart)
message("Rows in tumor.site: ", nrow(q.site.datamart))
print(v.script.log)
```

### Problems Encountered: `r nrow(v.script.log[!grepl("Success",v.script.log$log.action.taken),])`

```{r Error Log} 
message("Flagged Issues w/ Variables")
print(v.script.log[!grepl("Success",v.script.log$log.action.taken),])
```

### Descriptive Analysis

```{r Descriptive Analysis} 
for (tmp.col.index in 4:(ncol(q.site.datamart))) # Skip P#, E# and Site columns
  {  
  tmp.col.name <- names(q.site.datamart[tmp.col.index])                         
  BCRup.VisualizeDatamartColumn(q.site.datamart,tmp.col.name)
  BCRup.DescribeDatamartColumn(q.site.datamart, tmp.col.name)    
  }

# Provides a presentation similar to a standard 'Table One' 
for (tmp.col.index in 4:(ncol(q.site.datamart))) # Skip P#, E# and Site columns
  {  
  tmp.col.name <- names(q.site.datamart[tmp.col.index])
  BCRup.DescribeDatamartColumn(q.site.datamart, tmp.col.name)  
  }
```

### Export Encounters => Datamart

```{r Export Datamart}
message('Output directory: ',getwd())
v.filename <- paste0(v.output.dir,"BCDatamart-AllPts-AllEnctrs-",dataset$site,".csv")
write.csv(q.site.datamart, file = v.filename, row.names=FALSE)
message('Number of rows written: ',nrow(q.site.datamart))

v.filename <- paste0(v.output.dir,"BCDatamart-IssuesLog-",dataset$site,".csv")
write.csv(v.script.log[!grepl("Success",v.script.log$log.action.taken),], file = v.filename, row.names=FALSE)
message('Number of issues written: ',nrow(v.script.log[!grepl("Success",v.script.log$log.action.taken),]))
```

### Datamart Description: `r dataset$site`

```{r Datamart Description, echo=FALSE}
# use of 'print' in markdown creates a new bordered block in the output
print("Column Stats: gpc.stage.ss")
print(as.data.frame(table(q.site.datamart$gpc.stage.ss)))

print("Column Stats: NAACCR.0440.Grade.Descriptor")
print(as.data.frame(table(q.site.datamart$NAACCR.0440.Grade.Descriptor)))

message('Problems encountered:  ',nrow(v.script.log[!grepl("Success",v.script.log$log.action.taken),]))
print(paste0(Sys.time()))
v.script.log[!grepl("Success",v.script.log$log.action.taken),]
```

### Datamart Summary Statistics (ALL ELIGIBLE PATIENTS - Prior to consent): `r dataset$site`

```{r Datamart Summary Statistics - All Patients (Before), echo=FALSE}
message('GPC Site:                ', dataset$site)
message('Patients loaded:         ',length(unique(q.site.datamart$patient.num)))
message('Encounters exported:     ',nrow(q.site.datamart),' (num of rows written)')
message('Pts/Encounters eligible: ',length(q.site.datamart$gpc.enctr.eligible[q.site.datamart$gpc.enctr.eligible]))
message('Dx Age (Mean):           ',mean(unique(q.site.datamart[,c("patient.num","gpc.dx.age")])$gpc.dx.age))
message('GPC Birth Date (Min):    ',min(q.site.datamart$gpc.date.birth))
message('NAACCR Birth Date (Min): ',min(q.site.datamart$NAACCR.0240.Birth.Date))
message('GPC Birth Date (Max):    ',max(q.site.datamart$gpc.date.birth))
message('NAACCR Birth Date (Max): ',max(q.site.datamart$NAACCR.0240.Birth.Date))
message('GPC Dx Date (Min):       ',min(q.site.datamart$gpc.date.dx))
message('NAACCR Dx Date (Min):    ',min(q.site.datamart$NAACCR.0390.Dx.Date))
message('GPC Dx Date (Max):       ',max(q.site.datamart$gpc.date.dx))
message('NAACCR Dx Date (Max):    ',max(q.site.datamart$NAACCR.0390.Dx.Date))
```
