```{r input, echo=FALSE}
input <- source('dataset.R')$value

conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, submitter $name, $issues_other
dataset <- input$about

SITE <- dataset$site
```

Breast Cancer QA for `r SITE`
=============================

This report is generated using [bc_qa][] version`r system('hg id', intern=TRUE)`
For context, see [227].

[227]: https://informatics.gpcnetwork.org/trac/Project/ticket/227
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

*Load query terms and exclusion criteria from earlier article.*

```{r terms}
load("bc_terms_results.RData")
```


```{r txform, echo=FALSE}
# Transformation utilities:
source('bc_qa_txform.R')
```


```{r wrap, echo=FALSE}
# Don't wrap so much
options(width=300)
```

## Data Summary for `r SITE`

```{r about_dataset, echo=FALSE}
dataset[, c('record_id', 'site', 'name', 'bc_db', 'content_length')]
```

Issues noted by breast cancer research team via redcap:

```{r issues, echo=FALSE}
cat(dataset$issues_other)
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Patients, Encounters, and Observations per Variable",
  builder.summary(conn.site)[, c('variable', 'pat_qty', 'enc_qty', 'fact_qty')]
  )
```

## Terms in dataset from `r SITE`

### Variable Terms: Aligned and Misaligned

```{r, echo=FALSE}
v <- dbGetQuery(conn.site,
"
select id, name, concept_path
from variable
")

# Some sites have really long variable names
v$name <- substr(v$name, 1, 40)
v.ok <- v$concept_path %in% bcterm$term204$concept_path
v.good <- v[v.ok, ]
v.bad <- v[! v.ok, ]

data.frame(
  qty=nrow(v),
  qty.good=nrow(v.good),
  qty.bad=nrow(v.bad))
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Aligned Query Terms",
  v.good
  )
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Mis-Aligned Query Terms",
  v.bad
  )
```


### Fact Terms: Misaligned

For facts with concept paths that don't match any of the requested terms, how many patients, encounters, and facts?

```{r echo=FALSE}

# KLUDGE rather than adding a table... or even a temp table...
sql.qterm.table <- (
  function() {
    select <- lapply(bcterm$term204$concept_path,
                     function(p) paste0("select '", p, "' concept_path"))
    paste(select, collapse=' union all\n')
    })()

bad.concept.path <-
  dbGetQuery(conn.site,
             paste(
             "select count(distinct patient_num) qty_pat,
                     count(distinct encounter_num) qty_enc,
                     count(*) qty_fact,
                     substr(cd.name_char, 1, 30) name_char,
                     cd.concept_path
             from observation_fact f
             join concept_dimension cd
             on cd.concept_cd = f.concept_cd
             where not exists (select t204.concept_path from
             (", sql.qterm.table, ") t204
             where cd.concept_path like (t204.concept_path || '%'))
             group by cd.concept_path
             "))
```

```{r echo=FALSE, results='asis'}
ht(
  bad.concept.path,
  caption="Facts not aligned to Query Terms"
  )
```

## Analytic Dataset: One Record per Tumor

```{r}
tumor.site <- dbGetQuery(conn.site,
                         "select distinct encounter_num,
                         patient_num from observation_fact f
                         join concept_dimension cd
                         on cd.concept_cd = f.concept_cd
                         where cd.concept_path like '%naaccr%'")

bcterm$bc.dx.path
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$bc.dx.path, 'bc.dx',
                       get.var=v.enc)
```

We find `r nrow(tumor.site)` NAACCR-related encounters, `r length(which(!is.na(tumor.site$bc.dx)))` of which have breast cancer diagnoses; let's look at the first few:

```{r}
head(tumor.site)
```



### Sex, Age

Let's be careful to continue to rely only on paths. For nominal variables, we focus on segments of the concept path that distinguish it from the variable. For date-valued variables, we use the i2b2 `start_date`.

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['sex', 'concept_path'], 'sex')

tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['date.birth', 'concept_path'], 'date.birth',
                       get.var=v.enc)

head(tumor.site)
```

This function summarizes a nominal (enumerated) variable:

```{r}
enum.sum <- function(data, v,
                     key='patient_num') {
  summary(unique(data[, c(key, v)])[[v]])
}

enum.sum(tumor.site, 'sex')
```

Let's visualize the breakdown:

```{r}
prop <- enum.sum(tumor.site, 'sex')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```


```{r}
if (any(!is.na(tumor.site$date.birth))) {
  age.by.pat <- unique(survey.sample[, c('patient_num', 'age')])
  hist(age.by.pat$age)
  
  age.at.dx <- age.in.years(tumor.site$date.birth,
                            as.of=as.POSIXct(tumor.site$bc.dx))
  hist(age.at.dx)
  }
```

### Vital Status

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['vital.tr', 'concept_path'], 'vital.tr')
head(tumor.site)
```

To match between the tumor registry and the EHR, we can only join on patient num.

```{r}
tumor.site <- with.var.pat(tumor.site, conn.site,
                           bcterm$excl.all['deceased.ehr', 'concept_path'],
                           'deceased.ehr')
tumor.site <- with.var.pat(tumor.site, conn.site,
                           bcterm$excl.all['deceased.ssa', 'concept_path'],
                           'deceased.ssa')
tumor.site$vital <- vital.combine(tumor.site)

head(tumor.site)
```

#### Vital Status: Tumor Registry

```{r}
prop <- enum.sum(tumor.site, 'vital.tr')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

#### Vital Status: Deceased (from EHR)

```{r}
prop <- enum.sum(tumor.site, 'deceased.ehr')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

#### Vital Status: Deceased (from SSA)

```{r}
prop <- enum.sum(tumor.site, 'deceased.ssa')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

#### Vital Status: All Sources

```{r}
prop <- enum.sum(tumor.site, 'vital')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

### Language *TODO*

### Demographic Eligibility

```{r}
survey.sample <- check.demographics(tumor.site)
table(unique(survey.sample[, c('patient_num', 'female', 'adult', 'not.dead')])[, -1])
```
### Diagnostic Confirmation

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['confirm', 'concept_path'], 'confirm')

head(tumor.site[, c(-3:-6)])

prop <- enum.sum(tumor.site, 'confirm')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```


### Morphology Code

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['morphology', 'concept_path'], 'morphology')

head(tumor.site[, c(-3:-7)])

prop <- enum.sum(tumor.site, 'morphology')
pie(prop)
```

Those codes aren't very satisfying; let's put the names back on:

```{r echo=FALSE}
# Put the names back on
enum.name <- function(x, path.var) {
  df <- data.frame(
    concept_path=paste0(path.var,
                        # chop off 1st character
                        substr(x, 2, length(x))))
  
  df <- merge(df,
              dbGetQuery(conn.site,
                         'select name_char, concept_path from concept_dimension'),
              by='concept_path',
              all.x=TRUE)
  df$name_char
}
```

```{r}
prop <- data.frame(qty=prop, row.names=names(prop))
prop$name <- enum.name(row.names(prop), bcterm$excl['morphology', 'concept_path'])
prop <- prop[prop$qty> 0, ]
pie(prop$qty, labels=prop$name)
prop[, c('name', 'qty')]
```


#### Only tumor morphology was lobular carcinoma in situ?

```{r}
survey.sample$other.morph <- survey.sample$patient_num %in% check.morph(tumor.site, '8520/2')
```

### Stage

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['stage.ss', 'concept_path'], 'stage.ss')
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['stage.ajcc', 'concept_path'], 'stage.ajcc')
tumor.site$stage <- stage.combine(tumor.site)

head(tumor.site[, c(-3:-8)])
```

#### Stage: 3020 Derived SS2000

```{r}
prop <- enum.sum(tumor.site, 'stage.ss')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

#### Stage: 3430 Derived AJCC-7

```{r}
prop <- enum.sum(tumor.site, 'stage.ajcc')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

#### Stage: Combined

In [AJCC][], codes
 - 0xx      are stage 0
 - 1xx, 2xx are stage I
 - 3xx, 4xx are stage II
 - 5xx, 6xx are stage III
 - 7xx      are stage IV

[AJCC]: http://www.ccrcal.org/dsqc_pubs/v3_2010_forward/Data_Items/Derived_AJCC-7_Stage_Grp.htm

In [SS2000][], 7 is D Distant; per wikipedia, that's metastatic, i.e. stage 4.

[SS2000]: http://www.ccrcal.org/dsqc_pubs/v3_2010_forward/Data_Items/Derived_SS2000.htm

```{r}
prop <- enum.sum(tumor.site, 'stage')
pie(prop)
data.frame(qty=prop, row.names=names(prop))
```

### Prior Diagnosis: Sequence Number

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['seq.no', 'concept_path'], 'seq.no')
head(tumor.site[, c(-3:-12)])

prop <- enum.sum(tumor.site, 'seq.no')
pie(prop)
data.frame(qty=prop, row.names=names(prop))

```

## Summary and Issues for `r SITE`

```{r echo=FALSE, results='asis'}
ht(
  caption='Summary of exclusion criteria data',
  summary(tumor.site[, row.names(bcterm$excl)]),
  NA.string=''
  )
```

### Eligible Cohort from `r SITE`

```{r echo=FALSE}
survey.sample <- check.cases(tumor.site)
survey.sample.size <- count.cases(survey.sample)
```

```{r echo=FALSE, results='asis'}
ht(
  survey.sample.size
  )
```

### Issues for `r SITE`

Mis-aligned query terms: `r nrow(v.bad)`.

Mis-aligned facts: `r if(nrow(bad.concept.path) > 0) { sum(bad.concept.path$qty_fact) } else { 0 }`.

```{r}
show.issues(tumor.site, bcterm$excl)
```
