Breast Cancer QA: Exclusion Criteria from All Sites
===================================================

*Transformation and access utilities:*

```{r}
source('bc_qa_txform.R')
```


```{r echo=FALSE}
# Don't wrap so much
options(width=300) 
```


Let's summarize exclusion criteria for the current data sets:

*Restore work saved from earlier sections.*

```{r}
load("bc_data_excl.RData")
```

```{r}
current[c('record_id', 'site', 'name', 'filename')]
```


The framework looks like this:

```{r}
stats.empty <- function() {
  empty <- rep(0, nrow(var.excl) + 1)
  df <- data.frame(lapply(current$site, function(x) empty))
  names(df) <- current$site
  row.names(df) <- c('BC DX', row.names(var.excl))
  df
}

stats.empty()
```

Our data set has one row per case (tumor) and columns for encounter number, patient number, and our exclusion criteria. We're counting the number of distinct patient numbers where the some value of the exclusion criterion is provided in the data set.


```{r}
pat.with <- function(df, col) {
  ok.cases <- !is.na(df[[col]])
  length(unique(df$patient_num[ok.cases]))
}
```

Now let's crunch the data:

```{r}
excl.sites <- function(df) {
  for (ix in 1:nrow(current)) {
    site <- current$site[ix]
    message('working on site: ', ix, ' ', site)
    tumor.site <- bc.exclusions(site.data(site, dataDir, current))
    stats <- lapply(row.names(var.excl), function(n) pat.with(tumor.site, n))
    stats <- c(length(unique(tumor.site$patient_num)), stats)
    x <- unlist(stats)
    df[, site] <- x
  }
  df
}

stats <- excl.sites(stats.empty())
stats
```

For comparison with Vince's work, we can look at it the other way:

```{r}
stats.by.site <- t(stats)
stats.by.site <- stats.by.site[order(row.names(stats.by.site)), ]
stats.by.site
```


## Appendix: Query Terms


The query was sent to GPC sites 12/23/2014:

  - [GPC Honest Brokers: Time to run the breast cancer survey query][23dec]
    - Tamara McMahon Tue Dec 23 16:17:23 CST 2014 

[23dec]: http://listserv.kumc.edu/pipermail/gpc-honest-brokers/2014-December/000002.html

cf. [ticket:204][204]

[204]: https://informatics.gpcnetwork.org/trac/Project/ticket/204

From a KUMC data file, we extracted the variable table in [bc-variable.csv][222b] on Feb 9 using:

    sqlite3 -header -csv kumcBC.db 'select * from variable' >bc-variable.csv 

[222b]:  https://informatics.gpcnetwork.org/trac/Project/attachment/ticket/222/bc-variable.csv

*Note: R doubles backslashes when it displays strings.*

```{r}
var204 <- read.csv('bc-variable.csv',
                   stringsAsFactors=FALSE)

var204$name <- strip.counts(var204$name_char)
var204 <- var204[, c('id', 'concept_path', 'name')]
var204[, c('concept_path', 'name')]

```

We seem to have some duplicates:
```{r}
var204[duplicated(var204$concept_path), ]
data.frame(
  term.qty=nrow(var204),
  term.uniq.qty=length(unique(var204$concept_path))
  )
```

The first 4 exclusion criteria are taken directly from those terms:

```{r}
var.excl[, c('name', 'concept_path')]
```

The other two are a combination of 2 staging terms and 3 vital status terms:
```{r}
terms.aux <- data.frame(
  concept_path=c(
    '\\i2b2\\naaccr\\S:11 Stage/Prognostic Factors\\3020 Derived SS2000\\',
    '\\i2b2\\naaccr\\S:11 Stage/Prognostic Factors\\3430 Derived AJCC-7 Stage Grp\\',
    '\\i2b2\\naaccr\\S:4 Follow-up/Recurrence/Death\\1760 Vital Status\\',
    '\\i2b2\\Demographics\\Vital Status\\Deceased\\',
    '\\i2b2\\Demographics\\Vital Status\\Deceased per SSA\\'),
  name=c('stage.ss', 'stage.ajcc', 'vital.tr', 'deceased.ehr', 'deceased.ssa'))
terms.aux
```
