```{r echo=FALSE}
library(ggplot2)
library(reshape)
```

```{r input, echo=FALSE}
input <- source('dataset.R')$value

conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, submitter $name, $issues_other
dataset <- input$about

SITE <- dataset$site

version.bc_qa <- ''
try (
  version.bc_qa <- paste(' version', system('hg id', intern=TRUE)),
  silent=TRUE
  )

```

Breast Cancer QA for `r SITE`
=============================

**NOTE: Selected cohort and issues are summarized at the end of this report.**

This report is generated using [bc_qa][]`r version.bc_qa`.

For context, see [227].

[227]: https://informatics.gpcnetwork.org/trac/Project/ticket/227
[bc_qa]: https://bitbucket.org/gpcnetwork/bc_qa

*Load query terms and exclusion criteria from Query Terms and Exclusion Criteria article.*

```{r terms}
load("bc_terms_results.RData")
```


```{r txform, echo=FALSE}
# Transformation utilities:
source('bc_qa_txform.R')
```


```{r wrap, echo=FALSE}
# Don't wrap so much
options(width=300)
```

## Data Summary for `r SITE`

```{r about_dataset, echo=FALSE}
dataset[, c('record_id', 'site', 'name', 'bc_db', 'content_length', 'timestamp')]
```

Issues noted by breast cancer research team via redcap:

```{r issues, echo=FALSE}
cat(dataset$issues_other)
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Patients, Encounters, and Observations per Variable",
  builder.summary(conn.site)[, c('variable', 'pat_qty', 'enc_qty', 'fact_qty')]
  )
```

## Terms in dataset from `r SITE`

### Variable Terms: Aligned and Misaligned

```{r, echo=FALSE}
v <- dbGetQuery(conn.site,
paste0("
select id, name, concept_path, ", patch.umn('concept_path'), " patched_path
from variable
"))

# Some sites have really long variable names
v$name <- substr(v$name, 1, 40)
v.ok <- v$patched_path %in% bcterm$term204$concept_path
prefix.ok <-
  function(needle, haystack=bcterm$term204$concept_path)
    any(unlist(lapply(haystack,
                      function(h) grepl(h, needle, fixed=TRUE))))
v.close <- !v.ok
v.close[!v.ok] <- unlist(lapply(v$patched_path[!v.ok], prefix.ok))
           
v$patched_path[v$patched_path == v$concept_path] <- NA
v.good <- v[v.ok, ]
v.too.narrow <- v[v.close, ]
v.bad <- v[!v.ok & !v.close, ]

data.frame(
  qty=nrow(v),
  qty.good=nrow(v.good),
  qty.too.narrow=nrow(v.too.narrow),
  qty.bad=nrow(v.bad))
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Aligned Query Terms",
  v.good
  )
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Overly-specific Query Terms",
  v.too.narrow
  )
```

```{r echo=FALSE, results='asis'}
ht(
  caption="Mis-Aligned Query Terms",
  v.bad
  )
```


### Fact Terms: Misaligned

For facts with concept paths that don't match any of the requested terms, how many patients, encounters, and facts?

```{r echo=FALSE}

# KLUDGE rather than adding a table... or even a temp table...
sql.qterm.table <- (
  function() {
    select <- lapply(bcterm$term204$concept_path,
                     function(p) paste0("select '", p, "' concept_path"))
    paste(select, collapse=' union all\n')
    })()

bad.concept.path <-
  dbGetQuery(conn.site,
             paste(
             "select count(distinct patient_num) qty_pat,
                     count(distinct encounter_num) qty_enc,
                     count(*) qty_fact,
                     substr(cd.name_char, 1, 30) name_char,
                     cd.concept_path
             from observation_fact f
             join concept_dimension cd
             on cd.concept_cd = f.concept_cd
             where not exists (select t204.concept_path from
             (", sql.qterm.table, ") t204
             where", patch.umn('concept_path'), "
              like (t204.concept_path || '%'))
             group by cd.concept_path
             "))
```

```{r echo=FALSE, results='asis'}
if (nrow(bad.concept.path) > 0) {
  ht(
    bad.concept.path,
    caption="Facts not aligned to Query Terms"
    )
  } else {
    "No mis-aligned facts."
  }
```

## Analytic Dataset: Tumors, Inclusion Criteria

```{r}
tumor.site <- v.enc(conn.site,
                    subset(bcterm$term204,
                           # some sites don't have '0390 Date of Diagnosis'??
                           # Fall back to '0400 Primary Site'?
                           grepl('0390 Date of Diagnosis', concept_path)
                           )$concept_path,
                    'date.dx')

tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$t.incl[1,]$concept_path, 'seer.breast')
qty.seer.breast <- length(which(!is.na(tumor.site$seer.breast)))

tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$t.incl[2,]$concept_path, 'primary.site')
qty.primary.breast <- length(grep('^.C50', tumor.site$primary.site))
```

We find `r nrow(tumor.site)` tumors diagnosed, `r qty.primary.breast` of which have a primary site of breast cancer, and `r qty.seer.breast` are breast cancer by the SEER Site Recode.

Let's look at the first few. For nominal variables, note we **rely only on concept paths**; for conciseness, we focus on segments of the concept path that distinguish it from the variable's concept path. For date-valued variables, we use the i2b2 `start_date`.


```{r}
head(tumor.site)

tumor.site[duplicated(tumor.site$encounter_num), ]
```

```{r precompute, echo=FALSE}
tumor.site.pre <- bc.exclusions(conn.site)
survey.sample <- check.cases(tumor.site.pre)
survey.sample.size <- count.cases(survey.sample)
```

### Date of Diagnosis

How long ago were they diagnosed? (**Note: these are de-identified dates.**)

```{r}
if (any(!is.na(tumor.site$date.dx))) {
  hist(age.in.years(tumor.site$date.dx),
       main="Time since diagnosis",
       xlab="years")
  summary(age.in.years(tumor.site$date.dx))
  }
```


```{r}
if (any(!is.na(tumor.site$date.dx))) {
  recent <- subset(tumor.site, date.dx >=
        subset(bcterm$dx.date, txform == 'deid' & label == 'expanded')$start
      )
  hist(age.in.years(recent$date.dx),
       main="Time since diagnosis (Recent Only)",
       xlab="years")
  summary(age.in.years(recent$date.dx))
  }
```

**The Breast Cancer Cohort Characterization — Survey Sample report has `enddate <- as.POSIXct("2014-12-31")`. Why? The survey period ends 2014-05-01, and date shifting only goes backwards, no?**

How many patients with breast cancer diagnoses in the date range had 1 vs. more tumors?

```{r}
focus <- subset(survey.sample, bc.dx & recent.dx)
addmargins(table(as.data.frame(table(focus$patient_num))$Freq))
```

### Sex, Age

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['sex', 'concept_path'], 'sex')

tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['date.birth', 'concept_path'], 'date.birth',
                       get.var=v.enc)

head(tumor.site)
```

This function summarizes a nominal (enumerated) variable:

```{r}
enum.sum <- function(data, v,
                     key='patient_num') {
  table(unique(data[, c(key, v)])[[v]], useNA="ifany")
}

as.data.frame(addmargins(enum.sum(tumor.site, 'sex')))
```

Let's visualize the breakdown:

```{r}
prop <- enum.sum(tumor.site, 'sex')
pie(prop)
as.data.frame(addmargins(prop))
```


```{r}
if (any(!is.na(tumor.site$date.birth))) {
  x <- tumor.site[, c('encounter_num', 'patient_num', 'date.birth')]
  x$age <- age.in.years(x$date.birth)
  age.by.pat <- unique(x[, c('patient_num', 'age')])
  hist(age.by.pat$age)
  summary(age.by.pat$age)
  
  age.at.dx <- age.in.years(tumor.site$date.birth,
                            as.of=tumor.site$date.dx)
  hist(age.at.dx)
  }

if (any(!is.na(tumor.site$date.birth))) {
  summary(age.by.pat$age)
}
if (any(!is.na(tumor.site$date.birth))) {
  summary(age.at.dx)
}

```

### Vital Status

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['vital.tr', 'concept_path'], 'vital.tr')
head(tumor.site)
```

To match between the tumor registry and the EHR, we can only join on patient num.

```{r}
vital.agg <- mk.agg.by.pat()

tumor.site <- with.var.pat(tumor.site, conn.site,
                           bcterm$excl.all['vital.ehr', 'concept_path'],
                           'vital.ehr',
                           get.var=vital.agg)
tumor.site$vital <- vital.combine(tumor.site)

head(tumor.site)
```

#### Vital Status: Tumor Registry

```{r}
prop <- enum.sum(tumor.site, 'vital.tr')
pie(prop)
as.data.frame(addmargins(prop))
```

#### Vital Status: EHR

```{r}
prop <- enum.sum(tumor.site, 'vital.ehr')
pie(prop)
as.data.frame(addmargins(prop))
```

#### Vital Status: All Sources

```{r}
vital.per.pat <- unique(subset(tumor.site, select=c(patient_num, vital)))
prop <- table(vital.per.pat$vital, dnn=c('Alive'), useNA="ifany")
pie(prop)
as.data.frame(addmargins(prop))
```

#### Survival *TODO*

### Language

```{r}
lang.agg <- mk.agg.by.pat()

tumor.site <- with.var.pat(tumor.site, conn.site,
                           bcterm$excl.all['language', 'concept_path'],
                           'language',
                           get.var=lang.agg)

message('TODO: re-code UNKNOWN language as NA?')
```

```{r}
prop <- enum.sum(tumor.site, 'language')
pie(prop)
as.data.frame(addmargins(prop))
```

### Demographic Eligibility

```{r}
survey.sample <- check.demographics(tumor.site)

dem.states <- unique(survey.sample[, c('patient_num', 'female', 'adult', 'not.dead')])

# No person is both dead and alive etc., right?
stopifnot(
  length(which(duplicated(dem.states$patient_num)))
  == 0)

dem.states <- table(dem.states[, -1])
plot(dem.states,
     main="Demographic Eligibility")
dem.states
```

### Diagnostic Confirmation

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['confirm', 'concept_path'], 'confirm')

head(tumor.site[, c(-3:-6)])

prop <- enum.sum(tumor.site, 'confirm')
pie(prop)
as.data.frame(addmargins(prop))
```


### Morphology Code

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['morphology', 'concept_path'], 'morphology')

head(tumor.site[, c(-3:-7)])

prop <- enum.sum(tumor.site, 'morphology')
pie(prop)
```

Those codes aren't very satisfying; let's put the names back on:

```{r echo=FALSE}
# Find names of concepts whose paths end with tail
enum.name <- function(x) {
  names <- lapply(x, function(code) {
    name_char <- dbGetQuery(conn.site,
               "select distinct name_char
               from concept_dimension
               where concept_path like ('%' || :code || '%')",
               bind.data=data.frame(code=code))
    if (nrow(name_char) < 1) {
      '?'
    } else {
      name_char[1, 1]
    }
  })
  unlist(names)
}
```

```{r}
names(prop) <- enum.name(names(prop))
pie(prop)
as.data.frame(addmargins(prop))
```


#### Only tumor morphology was lobular carcinoma in situ?

```{r}
m <- excl.pat.morph(tumor.site, '8520/2')
x <- tumor.site[!m$ok & !is.na(m$ok), ]
x[order(x$patient_num, x$date.dx), c('patient_num', 'encounter_num', 'date.dx', 'primary.site', 'morphology')]
```

### Stage

```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['stage.ss', 'concept_path'], 'stage.ss')
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl.all['stage.ajcc', 'concept_path'], 'stage.ajcc')
tumor.site$stage.ajcc[grepl("999|900|888", tumor.site$stage.ajcc)] <- NA
tumor.site$stage.ss[grepl("9", tumor.site$stage.ss)] <- NA
tumor.site$stage <- stage.combine(tumor.site)

head(tumor.site[, c(-3:-8)])
```

#### Stage: 3020 Derived SS2000

```{r}
prop <- enum.sum(tumor.site, 'stage.ss')
pie(prop)
as.data.frame(addmargins(prop))
```

#### Stage: 3430 Derived AJCC-7

```{r}
prop <- enum.sum(tumor.site, 'stage.ajcc')
pie(prop)
as.data.frame(addmargins(prop))
```

#### Stage: Combined

In [AJCC][], codes
 - 0xx      are stage 0
 - 1xx, 2xx are stage I
 - 3xx, 4xx are stage II
 - 5xx, 6xx are stage III
 - 7xx      are stage IV

[AJCC]: http://www.ccrcal.org/dsqc_pubs/v3_2010_forward/Data_Items/Derived_AJCC-7_Stage_Grp.htm

In [SS2000][], 7 is D Distant; per wikipedia, that's metastatic, i.e. stage 4.

[SS2000]: http://www.ccrcal.org/dsqc_pubs/v3_2010_forward/Data_Items/Derived_SS2000.htm

```{r}
prop <- enum.sum(tumor.site, 'stage')
pie(prop)
as.data.frame(addmargins(prop))
```

Cross-table of AJCC7 and SS2000 (for in-date-range breast cancer tumors):

```{r}
survey.sample <- check.cases(tumor.site.pre)
stopifnot(all(tumor.site$patient_num == survey.sample$patient_num))
with(
  subset(tumor.site, survey.sample$bc.dx & survey.sample$recent.dx),
  addmargins(table(stage.ajcc, stage.ss, useNA="ifany"))
  )
```

### Prior Diagnosis: Sequence Number


```{r}
tumor.site <- with.var(tumor.site, conn.site,
                       bcterm$excl['seq.no', 'concept_path'], 'seq.no')
head(tumor.site[, c(-3:-12)])

prop <- enum.sum(tumor.site, 'seq.no')
pie(prop)
as.data.frame(addmargins(prop))

```

## Eligible Cohort

```{r echo=FALSE}
survey.sample <- check.cases(tumor.site)
survey.sample.size <- count.cases(survey.sample)
```

Looking at each exclusion criterion independently, how many patients and tumors does it leave?

```{r echo=FALSE, fig.width=12, fig.height=8}
style.dodge <- function(fig) {
  fig + geom_bar(stat = "identity", position="dodge") +
    labs(x="", y="") +
    theme(legend.position=c(0.1, 0.2)) +
    geom_text(data=fig$data, aes(label = result),
                  position = position_dodge(width=0.9), angle=45)
  }

x <- melt(cbind(survey.sample.size[c('ind.pat', 'ind.tumor'), ],
                c('patient','tumor')))
names(x)[1] <- 'kind'
names(x)[3] <- 'result'

style.dodge(
  ggplot(x, aes(x=variable, y=result, fill=kind))
  )

```

Now applying the exclusion critera successively, how many patients and tumors are left at each step?

```{r echo=FALSE, fig.width=12, fig.height=8}
x <- melt(cbind(survey.sample.size[c('cum.pat', 'cum.tumor'), ],
                c('patient','tumor')))
names(x)[1] <- 'kind'
names(x)[3] <- 'result'

style.dodge(
  ggplot(x, aes(x=variable, y=result, fill=kind))
  )
```

-----------

```{r}
crit <- subset(survey.sample, select=-c(encounter_num, patient_num, age))
not.excl <- reduce.logical(crit)
length(which(not.excl))
if (any(not.excl)) {
  tumor.ok <- tumor.site[not.excl, ]
  addmargins(table(tumor.ok$stage.ss, useNA="ifany"))
  }
```

## Summary and Issues for `r SITE`

```{r echo=FALSE, results='asis'}
ht(
  caption='Summary of exclusion criteria data',
  summary(tumor.site[, row.names(bcterm$excl)]),
  NA.string=''
  )
```

### Eligible Cohort from `r SITE`

```{r echo=FALSE, results='asis'}
ht(
  survey.sample.size[c('cum.pat'), ]
  )
```

### Issues for `r SITE`

Mis-aligned query terms: `r nrow(v.bad)`.

Mis-aligned facts: `r if(nrow(bad.concept.path) > 0) { sum(bad.concept.path$qty_fact) } else { 0 }`.

```{r echo=FALSE}
show.issues(tumor.site, bcterm$excl)
```
