Breast Cancer QA: Summary for All Sites
=======================================

Generated `r Sys.Date()`

```{r libraries}
library(RSQLite)
```

```{r previous results}
load("bc_fetch_results.RData")  # REDCap table listing of dataset file stats
load("bc_terms_results.RData")  # i2b2 BCcohort query terms
```

```{r shared transformation routines}
source('bc_qa_txform.R')  # Library of stuff
```


Summary of Submissions
======================

```{r fetch, echo=FALSE, results='asis'}
ht(fetch$dataset[, c('site', 'name', 'record_id', 'timestamp')])
```


Collect Exclusion Criteria Data Elements from Each Site
=======================================================

```{r site.code}
site.code <- factor(fetch$dataset$site)  # Provides factor of site code identifiers
```

```{r rbind.sites}
# move to bc_qa_txform.R?
# Produces dataframe which is 'union' of all site-specific datasets
#  - Input: site.tumor.data (list of site-specific dataframes)

rbind.sites <- function(df.site) {
  df <- NULL
  for (ix in 1:length(site.code)) {
    site <- site.code[ix]

    x <- df.site[[site]]

    if (nrow(x) > 0) {
      x$site <- site
      names <- names(x)
      if (is.null(df)) {
        df <- x
        } else {
          df <- rbind(df, x)
          }
      }
    }
  df
}
```

```{r tumor.all}
# Produces a list of site-specific dataframes less exclusions
site.tumor.data <- lapply(site.code, function(s) {
  message('Working on site: ', s)
  bc.exclusions(fetch$site.data(s))  # Produces site-specific results file
  })
names(site.tumor.data) <- levels(site.code)

```

Apply Exclusion Criteria
========================


```{r}
site.survey.sample <- lapply(site.code, function(s) {
  message('working on site: ', s)
  check.cases(site.tumor.data[[s]])
  })
names(site.survey.sample) <- levels(site.code)
site.survey.sample.size <- lapply(site.code, function(s) {
  count.cases(site.survey.sample[[s]])
  })

names(site.survey.sample.size) <- levels(site.code)
site.survey.sample.size  # Summary stats of exclusion criteria
```

Save for Later Sections
=======================

```{r}
# Summary stats for exclusion criteria
save(site.survey.sample.size, file="bc_excl_all_results.RData")
# Encounter-level TRUE/FALSE for exclusion criteria
save(site.survey.sample, file="tmp_sitesurveysample.RData")
# Source data
save(site.tumor.data, file="tmp_sitetumordata.RData")
```

Data Review for Descriptive Variables
=======================
- All patient data submitted is reviewed 
-- including patients excluded from the survey sample
-- this is consistent w/ method used to review eligibility variables

### Number of encounter rows by site

```{r}
bc.allsites.datamart <- rbind.sites(site.tumor.data)
table(bc.allsites.datamart$site)
```

### Patient & Encounter Counts
```{r}
bad <- list("")
bad$ttls <- setNames(aggregate(bc.allsites.datamart$patient_num,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","Pts"))
tmp.ttls  <- setNames(aggregate(bc.allsites.datamart$encounter_num,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","Enctrs"))
bad$ttls <- merge(bad$ttls,tmp.ttls)
bad$ttls
```

### Number of unique patients w/in each variable
```{r}
bad$ptcnts <- setNames(aggregate(bc.allsites.datamart$patient_num,
                                 by=list(bc.allsites.datamart$site),
                                 function(x) length(unique(x))),
                                 c("Site","Pts"))
tmp.stats <- setNames(aggregate(subset(bc.allsites.datamart,!is.na(er.csf.1))$patient_num,
                                 by=list(subset(bc.allsites.datamart,!is.na(er.csf.1))$site),
                                 function(x) length(unique(x))),
                                 c("Site","ER1"))
bad$ptcnts <- merge(bad$ptcnts,tmp.stats)
tmp.stats <- setNames(aggregate(subset(bc.allsites.datamart,!is.na(pr.csf.2))$patient_num,
                                 by=list(subset(bc.allsites.datamart,!is.na(er.csf.1))$site),
                                 function(x) length(unique(x))),
                                 c("Site","PR1"))
bad$ptcnts <- merge(bad$ptcnts,tmp.stats)
tmp.stats <- setNames(aggregate(subset(bc.allsites.datamart,!is.na(her2.csf.15))$patient_num,
                                 by=list(subset(bc.allsites.datamart,!is.na(her2.csf.15))$site),
                                 function(x) length(unique(x))),
                                 c("Site","HER2"))
bad$ptcnts <- merge(bad$ptcnts,tmp.stats)
tmp.stats <- setNames(aggregate(subset(bc.allsites.datamart,!is.na(mgs.method.csf.22))$patient_num,
                                 by=list(subset(bc.allsites.datamart,!is.na(mgs.method.csf.22))$site),
                                 function(x) length(unique(x))),
                                 c("Site","MGS-Method"))
bad$ptcnts <- merge(bad$ptcnts,tmp.stats)
tmp.stats <- setNames(aggregate(subset(bc.allsites.datamart,!is.na(mgs.score.csf.23))$patient_num,
                                 by=list(subset(bc.allsites.datamart,!is.na(mgs.score.csf.23))$site),
                                 function(x) length(unique(x))),
                                 c("Site","MGS Score"))
bad$ptcnts <- merge(bad$ptcnts,tmp.stats)

bad$ptcnts
```

### Distinct values

```{r}
tmp.stats  <- setNames(aggregate(bc.allsites.datamart$er.csf.1,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","ER1"))
bad$unique <- tmp.stats
tmp.stats  <- setNames(aggregate(bc.allsites.datamart$pr.csf.2,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","PR1"))
bad$unique <- merge(bad$unique,tmp.stats)
tmp.stats  <- setNames(aggregate(bc.allsites.datamart$her2.csf.15,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","HER2"))
bad$unique <- merge(bad$unique,tmp.stats)
tmp.stats  <- setNames(aggregate(bc.allsites.datamart$mgs.method.csf.22,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","MGS-Method"))
bad$unique <- merge(bad$unique,tmp.stats)
tmp.stats  <- setNames(aggregate(bc.allsites.datamart$mgs.score.csf.23,
                               by=list(bc.allsites.datamart$site),
                               function(x) length(unique(x))),
                               c("Site","MGS-Score"))
bad$unique <- merge(bad$unique,tmp.stats)

bad$unique
```
