Breast Cancer QA: Query Terms and Exclusion Criteria
====================================================

```{r, ECHO=FALSE}
load("bc_data_summary.RData")
```

## Query Terms

 Query sent to GPC sites 12/23/2014:

  - [GPC Honest Brokers: Time to run the breast cancer survey query][23dec]
    Tamara McMahon
    Tue Dec 23 16:17:23 CST 2014 

[23dec]: http://listserv.kumc.edu/pipermail/gpc-honest-brokers/2014-December/000002.html

cf. [ticket:204][204]

[204]: https://informatics.gpcnetwork.org/trac/Project/ticket/204

From a KUMC data file, we extracted the variable table in [bc-variable.csv][222b] on Feb 9 using:

    sqlite3 -header -csv kumcBC.db 'select * from variable' >bc-variable.csv 

[222b]:  https://informatics.gpcnetwork.org/trac/Project/attachment/ticket/222/bc-variable.csv

*Note: R doubles backslashes when it displays strings.*

```{r}
var204 <- read.csv('bc-variable.csv',
                   stringsAsFactors=FALSE)

var204$name <- strip.counts(var204$name_char)
var204 <- var204[, c('id', 'concept_path', 'name')]
head(var204, n=10)

```

## Analytic Dataset is based on Breast Cancer Diagnosis

### Encounter variable

For present/absent variables, we can capture the observation fact start date while we're at it using this query:

```{r}
v.enc <- function(conn, var.path, var.name) {
  sql.summary <- '
  select f.encounter_num, f.patient_num, f.start_date
  from observation_fact f
  join concept_dimension cd
  on cd.concept_cd = f.concept_cd
  join variable v
  on cd.concept_path like (v.concept_path || \'%\')
  where v.concept_path = ?
  '
  per.enc <- dbGetPreparedQuery(conn, sql.summary, bind.data=data.frame(path=var.path))
  per.enc$start_date <- as.POSIXct(per.enc$start_date)

  names(per.enc)[3] <- var.name
  per.enc
  }
```

Let's pick out the breast cancer term; oddly, it seems to occur in the query twice, so we use `unique()`:

```{r}
var.dx <- unique(var204[grep('Breast', var204$name), c('concept_path', 'name')])
stopifnot(nrow(var.dx) == 1)
var.dx[, c('name', 'concept_path')]
```

### Breast Cancer Diagnoses at a site

```{r}
conn.kumc <- site.data('KUMC', dataDir, current)

tumor.kumc <- v.enc(conn.kumc, var.dx$concept_path, 'bc.dx')
head(tumor.kumc)
```


## Exclusion Criteria

From Tamara McMahon's message of Thursday, February 19, 2015 9:03 AM:

> Here are the variables needed for exclusion (e.g., minimum variables required by all institutions)
 
```{r}
var.exclusion <- c(
       'Sex',
       'Sequence Number',
       'Diagnostic Confirmation',
       'Morphology Code',
       'Derived AJCC-7 Grp and/or SS2000',
       'Vital Status'
)
length(var.exclusion)
```


### Sex


```{r}
var.sex <- var204[grep('0220 Sex', var204$name), ]
stopifnot(nrow(var.sex) == 1)
var.sex
```

For nominal variables, we focus on segments of the concept path that distinguish it from the variable.

```{r}
v.enc.nominal <- function(conn, var.path, var.name) {
  sql.summary <- '
  select f.encounter_num, substr(cd.concept_path, length(v.concept_path)) tail
  from observation_fact f
  join concept_dimension cd
  on cd.concept_cd = f.concept_cd
  join variable v
  on cd.concept_path like (v.concept_path || \'%\')
  where v.concept_path = ?
  '
  per.enc <- dbGetPreparedQuery(conn, sql.summary, bind.data=data.frame(path=var.path))
  per.enc$tail <- as.factor(per.enc$tail)
  
  names(per.enc)[2] <- var.name
  per.enc
  }

```

```{r}
with.var <- function(data, conn, path, name,
                  get.var=v.enc.nominal) {
  merge(data, get.var(conn, path, name),
        by=c('encounter_num'))
}
```

```{r}
enum.sum <- function(data, v,
                     key='patient_num') {
  summary(unique(data[, c(key, v)])[[v]])
}
```

```{r}
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.sex$concept_path, 'sex')

head(tumor.kumc)

prop <- enum.sum(tumor.kumc, 'sex')
pie(prop)
prop

```


### Sequence Number

```{r}
var.seq <- var204[grep('0380 Sequence Number', var204$name), ]
stopifnot(nrow(var.seq) == 1)
var.seq
```


```{r}
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.seq$concept_path, 'seq.no')
head(tumor.kumc)

prop <- enum.sum(tumor.kumc, 'seq.no')
pie(prop)
prop

```

### Diagnostic Confirmation

```{r}
var.confirm <- var204[grep('0490 Diagnostic Confirmation', var204$name), ]
stopifnot(nrow(var.confirm) == 1)
var.confirm
```


```{r}
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.confirm$concept_path, 'confirm')

head(tumor.kumc)

prop <- enum.sum(tumor.kumc, 'confirm')
pie(prop)
prop
```

### Morphology Code

```{r}
var.morph <- var204[grep('0521 Morph', var204$name), ]
stopifnot(nrow(var.morph) == 1)
var.morph
```


```{r}
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.morph$concept_path, 'morphology')

head(tumor.kumc)

prop <- enum.sum(tumor.kumc, 'morphology')
pie(prop)
prop
```

### Stage

Derived AJCC-7 Grp and/or SS2000

```{r}
var.stage.ss <- var204[grep('3020 Derived SS2000', var204$name), ]
var.stage.ajcc <- var204[grep('AJCC-7', var204$name), ]

stopifnot(nrow(var.stage.ss) == 1)
stopifnot(nrow(var.stage.ajcc) == 1)
var.stage.ajcc
var.stage.ss
```


```{r}
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.stage.ss$concept_path, 'stage.ss')
tumor.kumc <- with.var(tumor.kumc, conn.kumc, var.stage.ajcc$concept_path, 'stage.ajcc')
tumor.kumc$stage <- as.factor(paste(as.character(tumor.kumc$stage.ss),
                                    as.character(tumor.kumc$stage.ajcc)))
  
head(tumor.kumc)

prop <- enum.sum(tumor.kumc, 'stage.ss')
pie(prop)
prop
prop <- enum.sum(tumor.kumc, 'stage.ajcc')
pie(prop)
prop
prop <- enum.sum(tumor.kumc, 'stage')
pie(prop)
prop
```
