Breast Cancer QA Part 1: Fetch Data Files
=========================================

```{r file_header, echo=FALSE}
doc <- list(
  run.date=date(),
  version=substr(system(' git rev-parse HEAD', intern=TRUE), 1, 16),
  copyright_year=2016,
  copyright_holder="University of Kansas Medical Center"
  )

# Don't wrap so much
options(width=300) 
```

_This is part of [GPC Breast Cancer QA][about],
copyright (c) `r doc$copyright_year` `r doc$copyright_holder`.
This program was run on `r doc$date` using [source code][src]
version `r doc$version`._

[src]: https://github.com/kumc-bmi/bc_qa
[about]: README.md

 - **inputs**:
   - Survey responses and data files submitted to the
     [GPC Breast Cancer Project 1: review of survey population][4044]
     REDCap project.
   - `config.csv` - one-record file of configuration settings
 - *modules / subprograms*
   - bc_access.py
 - **outputs**
   - local copies of data files
     - unpacked
   - `dataset.csv` with one row per non-obsolete submission,
     giving the filename, submitter, timestamp, etc.

[4044]: https://redcap.kumc.edu/redcap_v5.7.7/index.php?pid=4044
[pyenv]: http://docs.python-guide.org/en/latest/dev/virtualenvs/


## Configuration

This script relies on a `config.csv` file to specify a number of parameters:

  - url - REDCap API url
  - api_key - API key to read from the project
  - datafile_dir - where to put fetched files
  - python_env - [python virtual environment][pyenv] where libraries
    for `bc_access.py` can be found

```{r read_config, echo=TRUE, error=FALSE}
config <- read.csv('config.csv', stringsAsFactors=FALSE)

missing_config <- setdiff(c('url', 'api_key', 'datafile_dir', 'python_env'),
                          names(config))
stopifnot(length(missing_config) == 0)
```


```{r wrap_python, echo=FALSE}
# Wrap bc_access.py functions as a list of R functions

make_python_runner <- function(.system2, env,
                               virtual_environment, py_script)
  # Args:
  #   .system2: access to run subprograms (typically=system2)
  #   env: character vector of name=value bindings
  #   virtual_environment: directory containing python binaries and libraries
  #     ref: http://docs.python-guide.org/en/latest/dev/virtualenvs/
  #   py_script: full path to python script
  # returns: a function to run just that python script, given args,
  #        and return its output
  function(args, input=NULL) {
    python_with_libraries <- file.path(virtual_environment, 'bin', 'python')
    print('run python:')
    print(c(python_with_libraries, py_script, args))
    .system2(python_with_libraries, c(py_script, args),
             stdout=TRUE, env=env, input=input)
    }

make_bc_access = function(.system2, virtual_environment, src_dir, api_key) {
  run_bc_access <- make_python_runner(.system2,
                                      env=paste0('api_key=', api_key),
                                      virtual_environment=virtual_environment,
                                      py_script=file.path(src_dir, 'bc_access.py'))
  list(
    export=function() run_bc_access(args=c('--key', 'api_key', 'export')),
    fetch=function() run_bc_access(args=c('--key', 'api_key', 'fetch')),
    normalize=function(input) run_bc_access(args=c('normalize'), input=input)
    )
}

```

The `make_bc_access` function encapsulates our python subprogram as
a list of R functions:

```{r make_bc_access}
bc_qa_files <- getwd()

bc_access <- make_bc_access(.system2=system2,
                            virtual_environment=config$python_env,
                            src_dir=bc_qa_files,
                            api_key=config$api_key)
```


Let's fetch all the form data submitted to the REDCap project:

```{r fetch_form_data, echo=TRUE}
parse_csv <- function(text)
  read.csv(textConnection(text),
           stringsAsFactors=FALSE)

fetch <- list()
fetch$form <- parse_csv(bc_access$export())
fetch$form[, c('record_id', 'institution', 'name', 'obsolete')]
```


Next, we fetch the attached data files:
 
```{r fetch_data_files, echo=TRUE}
# bc_access writes data files to the current directory,
# so change directories before we run it
setwd(config$datafile_dir)
# get rid of existing data files?
# system("rm *")
fetch$fetch.detail <- bc_access$fetch()
setwd(bc_qa_files)  # come back
```


Many of the files are compressed in various ways; uncompress them and
prefix them with site identifier and REDCap record id:

```{r normalize}
setwd(config$datafile_dir)
fetch$file <- parse_csv(bc_access$normalize(fetch$fetch.detail))
setwd(bc_qa_files)  # come back
fetch$file
```

Let's combine the info about the submitted datasets:

```{r merge_dataset_info}
dataset <- merge(fetch$file, fetch$form,
                       by='record_id')
dataset <- dataset[order(dataset$site),]
# 'gpc_breast_cancer_deidentified_data_timestamp' is unwieldy
names(dataset)[grep('timestamp', names(dataset))] <- 'timestamp'
fetch$dataset <- dataset
fetch$dataset[, c('record_id', 'timestamp', 'site', 'name', 'bc_db', 'patient_qty')]
```

Save data about submissions:

```{r save}
write.csv(fetch$dataset[, c('record_id', 'timestamp', 'site', 'name', 'bc_db')],
          file='dataset.csv')
```
