Breast Cancer QA: Query Terms and Exclusion Criteria
====================================================

```{r, echo=FALSE}
# Don't wrap so much
options(width=300) 
```

*Transformation utilities:*

```{r}
source('bc_qa_txform.R')
```


Query sent to GPC sites 12/23/2014:

  - [GPC Honest Brokers: Time to run the breast cancer survey query][23dec]
    Tamara McMahon
    Tue Dec 23 16:17:23 CST 2014 

[23dec]: http://listserv.kumc.edu/pipermail/gpc-honest-brokers/2014-December/000002.html

As noted in [ticket:204][204], the query is also on babel.

[204]: https://informatics.gpcnetwork.org/trac/Project/ticket/204

From a KUMC data file, we extracted the variable table using:

    sqlite3 -header -csv kumcBC.db 'select * from variable' >bc-variable.csv 

and attached it to #204 as [bc-variable.csv][222b] on Feb 9.

[222b]:  https://informatics.gpcnetwork.org/trac/Project/attachment/ticket/222/bc-variable.csv

*Note: R doubles backslashes when it displays strings.*

```{r}
bcterm <- list()

term204 <- read.csv('bc-variable.csv',
                    stringsAsFactors=FALSE)

term204$name <- strip.counts(term204$name_char)
term204 <- term204[, c('id', 'concept_path', 'name')]
```

We seem to have some duplicates; let's eliminate them:
```{r}
term204[duplicated(term204$concept_path), ]
data.frame(
  term.qty=nrow(term204),
  term.uniq.qty=length(unique(term204$concept_path))
  )
term204 <- term204[!duplicated(term204$concept_path), ]
```

These are the `r nrow(term204)` query terms:

```{r}
term204$naaccr.code <- substr(term204$name, 1, 4)
term204[, c('concept_path', 'name')]
bcterm$term204 <- term204
```


## Exclusion Criteria

**TODO: trace these to a more authoritative source**

From Tamara McMahon's message of Thursday, February 19, 2015 9:03 AM:

> Here are the variables needed for exclusion (e.g., minimum variables required by all institutions)
>  - Sex
>  - Sequence Number
>  - Diagnostic Confirmation
>  - Morphology Code
>  - Derived AJCC-7 Grp and/or SS2000
>  - Vital Status

Note that stage (AJCC, SS2000) and Vital Status are combinations of terms:

```{r}
t.excl <- data.frame(
  naaccr.code=c('0220',
                '0380',
                '0490',
                '0521',
                '3020',
                '3430',
                '1760'),
  stringsAsFactors=FALSE
  )

t.excl <- merge(t.excl, term204[, c('naaccr.code', 'name', 'concept_path')],
                by='naaccr.code', all.x=TRUE)

deceased.ehr <- merge(
  data.frame(
  concept_path=c(
    '\\i2b2\\Demographics\\Vital Status\\Deceased\\',
    '\\i2b2\\Demographics\\Vital Status\\Deceased per SSA\\'
    )),
  term204[, c('name', 'concept_path')],
  by='concept_path')
deceased.ehr$naaccr.code <- NA

combo <- data.frame(
  name=c('stage', 'vital'),
  naaccr.code=NA,
  concept_path=NA)
t.excl <- rbind(t.excl, deceased.ehr, combo)

row.names(t.excl) <- c('sex', 'seq.no', 'confirm', 'morphology',
                       'stage.ss', 'stage.ajcc',
                       'vital.tr', 'deceased.ehr', 'deceased.ssa',
                       'stage', 'vital')
bcterm$excl.all <- t.excl
bcterm$excl.all

bcterm$excl <- t.excl[c('sex', 'seq.no', 'confirm', 'morphology',
                       'stage', 'vital'), ]
bcterm$excl
```


## Save For Next Section

```{r}
save(bcterm, file="bc_terms_results.RData")
```
