Breast Cancer QA Part 2: Site Data Summary
==========================================

Recall the results of fetching the files:

```{r}
load("bc_download.RData")
dataDir
names(got)

submission$record_id <- as.numeric(row.names(submission))
current <- submission[submission$obsolete == '', ]
nrow(current)
```

The following `r nrow(current)` records are current:

```{r}
current[c('record_id', 'name', 'institution')]
```

## Connecting File Info with CRF Info

```{r}
fileinfo <- (function() {
  filedesc <- grep('patients', got$normalizelog, value=TRUE)
  m <- regexec('[^:]+:[^:]+:(([^-]+)-([0-9]+)[^ ]+) has ([0-9]+) patients.', filedesc)
  lm <- regmatches(filedesc, m)
  df <- do.call(rbind.data.frame, lm)
  names(df) <- c('line', 'filename', 'site', 'record_id', 'n.patient')
  row.names(df) <- df$record_id
  df$record_id <- as.numeric(as.character(df$record_id))
  df[order(df$record_id), -1]
  })()
fileinfo
```

Now we can merge the file info with the redcap data:

```{r}
current$filename <- fileinfo$filename
current$site <- fileinfo$site
current$n.patient <- fileinfo$n.patient
current[, c('site', 'n.patient', 'filename')]
```


### SQL Queryies on Site Data

```{r echo=FALSE}
site.data <- function(target, dataDir, current) {
  library(RSQLite)

  f <- file.path(dataDir, subset(current, site == target)$filename)
  dbConnect(SQLite(), dbname=f)
}
```

For example:

```{r}
conn.kumc <- site.data('KUMC', dataDir, current)
dbGetQuery(conn.kumc, 'select count(*) from patient_dimension')
```


### Site Data Summary

This is the [data summary query from BuilderSaga][sqlsum], tweaked to add encounter qty:

[sqlsum]: https://informatics.gpcnetwork.org/trac/Project/wiki/BuilderSaga#DataSummary

```{r}
builder.summary <- function(conn) {
  sql.summary <- '
  select v.concept_path, v.name_char,
  count(distinct patient_num) pat_qty, count(distinct encounter_num) enc_qty, count(*) fact_qty
  from observation_fact f
  join concept_dimension cd
  on cd.concept_cd = f.concept_cd
  join variable v
  on cd.concept_path like (v.concept_path || \'%\')
  group by v.concept_path, v.name_char
  '
  per.var <- dbGetQuery(conn, sql.summary)
  per.var$variable <- strip.counts(per.var$name_char)
  per.var
  }

strip.counts <- function(text) {
  gsub('\\[.*', '', text)
}
```

```{r}
conn.kumc <- site.data('KUMC', dataDir, current)

builder.summary(conn.kumc)[, c('variable', 'pat_qty', 'enc_qty', 'fact_qty')]
```


## Save For Next Section

```{r}
save(dataDir, got, submission,
     current, site.data, builder.summary, strip.counts,
     file="bc_data_summary.RData")
```
