Breast Cancer QA: Query Terms and Exclusion Criteria
====================================================

```{r, echo=FALSE}
# Don't wrap so much
options(width=300) 
```

*Transformation utilities:*

```{r}
source('bc_qa_txform.R')
```


Query sent to GPC sites 12/23/2014:

  - [GPC Honest Brokers: Time to run the breast cancer survey query][23dec]  
    Tamara McMahon  
    Tue Dec 23 16:17:23 CST 2014 

[23dec]: http://listserv.kumc.edu/pipermail/gpc-honest-brokers/2014-December/000002.html

As noted in [ticket:204][204], the query is on babel, where a few revisions have been incorporated. 
`bc-variable.csv` is derived from the XML query definition via `qterms.py`.
*Then 0560 Sequence Number–Hospital was added to `bc-variable.csv` manually, since it's not part
of the ontology on babel. See [185][]*

[204]: https://informatics.gpcnetwork.org/trac/Project/ticket/204
[185]: https://informatics.gpcnetwork.org/trac/Project/ticket/185


```{r}
bcterm <- list()

term204 <- read.csv('bc-variable.csv',
                    stringsAsFactors=FALSE)

term204$name <- strip.counts(term204$name_char)
term204 <- term204[, c('id', 'concept_path', 'name')]

hlevel <- function(path,
                   sep='\\') {
  segments <- strsplit(path, sep, fixed=TRUE)
  unlist(lapply(segments, length))
}

term204$naaccr.code <- ifelse(
  # Is it in a NAACCR section?
  grepl('^.i2b2.naaccr.S:[0-9]+', term204$concept_path) &
  # Is it at item level?
  hlevel(term204$concept_path) == 5,
    # Yes. Get item number.
  substr(term204$name, 1, 4),
    # No.
  NA)

# Prune initial \i2b2\ segment, which gives no information.
# See also ticket #201
# https://informatics.gpcnetwork.org/trac/Project/ticket/201
term204$concept_path <- gsub('^\\\\i2b2', '', term204$concept_path)
```

*We seem to have `r length(which(duplicated(term204$concept_path)))` duplicates:*

```{r echo=FALSE, results='asis'}
ht(
  term204[duplicated(term204$concept_path), ]
  )
```

Let's eliminate them:

```{r}
data.frame(
  term.qty=nrow(term204),
  term.uniq.qty=length(unique(term204$concept_path))
  )
term204 <- term204[!duplicated(term204$concept_path), ]

bcterm$term204 <- term204
```

These are the `r nrow(term204)` query terms:

```{r, echo=FALSE, results='asis'}

ht(
  term204[, c('concept_path', 'name')]
  )
```

### Alignment by "patching"


We can accomodate analagous paths to a certain extent; e.g. the "Abridged" NAACCR hierarchy:

```{r echo=FALSE}
bcterm$fixes <- read.csv(textConnection(
    'from,to,note
\\i2b2,,initial segment matching is tedious
\\I2B2,,initial segment matching is tedious
\\Abridged\\Demographics,\\naaccr\\S:2 Demographic,Abridged
\\Cancer Cases\\0,\\naaccr\\S:,UMN capitalization and section markup
\\Cancer Cases\\1,\\naaccr\\S:1,UMN
\\Cancer Cases\\SEER Site Summary,\\naaccr\\SEER Site,UMN
Type and Behav ICD-O-3,Type&Behav ICD-O-3,i2b2 ampersand-in-XML work-arounds
Type/Behav ICD-O-3,Type&Behav ICD-O-3,ampersand
S:1 Cancer Identification\\0560 Hospital Sequence Number,S:6 Hospital-Specific\\0560 Sequence Number--Hospital,UMN
2869 CS Site-Specific Factor 15,2869 CS Site-Specific Factor15,UMN
2876 CS Site-Specific Factor 22,2876 CS Site-Specific Factor22,UMN             
2877 CS Site-Specific Factor 23,2877 CS Site-Specific Factor23,UMN
'))
```

```{r}
bcterm$fixes

library(RSQLite)
test.patch <- function(before, expected) {
  conn <- dbConnect(SQLite(), dbname=':memory:')
  actual <- dbGetPreparedQuery(conn, paste("select", patch.umn(':x'), 'patched_path'),
                               bind.data=data.frame(x=before))
  #message("expected:", expected)
  #message("  actual:", actual)
  stopifnot(actual == expected)
  }

test.patch('\\I2B2\\Cancer Cases\\02 Demographic\\0220 Sex\\',
           '\\naaccr\\S:2 Demographic\\0220 Sex\\')
test.patch('\\I2B2\\Cancer Cases\\11 Stage/Prognostic Factors\\2876 CS Site-Specific Factor 22\\',
           '\\naaccr\\S:11 Stage/Prognostic Factors\\2876 CS Site-Specific Factor22\\')
test.patch('\\Cancer Cases\\01 Cancer Identification\\0560 Hospital Sequence Number\\',
           '\\naaccr\\S:6 Hospital-Specific\\0560 Sequence Number--Hospital\\')


```

## Inclusion Criteria: Recent Breast Cancer Diagnosis

The SEER definition of breast cancer is defined as the relevant primary site excluding
certain histologies. If the pre-computed combination is not available, we can compute it.

```{r}
t.incl <- term204[
  grepl('SEER Site\\Breast', term204$concept_path, fixed=TRUE) |
  grepl('0400 Primary Site', term204$concept_path, fixed=TRUE),
  c('naaccr.code', 'name', 'concept_path')]
rownames(t.incl) <- c('seer.breast', 'primary.site')
t.incl
bcterm$t.incl <- t.incl
```

```{r}
dx.date <- read.csv(textConnection('
txform,label,start,end
id,original,2013-05-01,2014-05-01
deid,original,2012-05-01,2014-05-01
id,expanded,2013-01-01,2014-05-01
deid,expanded,2012-01-01,2014-05-01'))
dx.date$start <- as.Date(dx.date$start)
dx.date$end <- as.Date(dx.date$end)

dx.date

bcterm$dx.date <- dx.date
```

```{r eval=FALSE, echo=FALSE}
#@@ TODO http://www.r-fiddle.org/#/fiddle?id=e78ubsxJ&version=1
library(ggplot2)
library(reshape)

ggplot(dx.date,
       aes(x = factor(paste(txform, label)),
           ymin=start, ymax=end, y=start,
           fill=factor(paste(txform, label))
           ),
       flatten=0) +
  geom_bar(stat = "identity") +
  coord_cartesian(ylim=c(as.Date('2010-01-01'),
                         as.Date('2015-01-01'))) +
  theme(legend.position = "none",
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

## Exclusion Criteria

In [Share Thoughts on Breast Cancer Study][18Dec] GPC Global Webinar, December 18, 2014, on Slide 8 *Selection Criteria*:

[18Dec]: http://listserv.kumc.edu/pipermail/gpc-all/attachments/20141215/f655d41b/attachment-0001.pptx

> Exclude from the SURVEY sample if:
>  - Sex not equal to female
>  - Less than 18 years of age
>  - Prior cancer diagnosis
>  - Breast cancer was not microscopically confirmed
>  - Only tumor morphology was lobular carcinoma in situ
>  - Stage IV breast cancer
>  - Known to be deceased
>  - Non-English speaking (for now)

*The BC team agreed 4 Mar to leave the Non-English constraint aside at this stage.*

Note that stage (AJCC, SS2000), sequence number, and Vital Status are combinations of terms:

```{r echo=FALSE}
t.excl <- rbind(
  merge(
    data.frame(
      naaccr.code=c('0220',
                    '0240',
                    '1760',
                    '0490',
                    '0521',
                    '3020',
                    '3430',
                    '0380',
                    '0560')),
    term204,
    sort=FALSE
    )[, c('naaccr.code', 'name', 'concept_path')],
  
  merge(
    data.frame(
      concept_path=c(
        '\\Demographics\\Vital Status\\'
        )),
    term204,
    sort=FALSE
    )[, c('naaccr.code', 'name', 'concept_path')],
  
  data.frame(
    naaccr.code=NA,
    name=c('stage', 'seq.no', 'vital'),
    concept_path=NA)
  )

row.names(t.excl) <- c('sex', 'date.birth',
                       'vital.tr',
                       'confirm', 'morphology',
                       'stage.ss', 'stage.ajcc',
                       'seq.no.380', 'seq.no.560',
                       'vital.ehr',
                       'stage', 'seq.no', 'vital')

bcterm$excl.all <- t.excl

bcterm$excl <- t.excl[c('sex', 'date.birth', 'vital',
                        'confirm', 'morphology', 'stage', 'seq.no'), ]
```

```{r echo=FALSE, results='asis'}
ht(caption="Exclusion Criteria Terms (All)",
   bcterm$excl.all)
ht(caption="Exclusion Criteria Terms (Primary)",
   bcterm$excl)
```

## Save For Next Section

```{r}
save(bcterm, file="bc_terms_results.RData")
```
