# Breast Cancer - Roll-up Site Data
```{r echo=FALSE}
# ============================================================================
###  Creates site-specific tumor-level datamart
# Input Dataframes
#   dataset       - desriptors for site's input dataset 
#   tumor.site    - pt tumors listing
#   survey.sample - complex frame created by bc_excl
# 
# Note - script was modeled after 'bc_random_smaple.rmd'
#
# 11-Sep Genesis
# 22-Sep SAVEPOINT (RMD file committed to TortoiseHg)
# ============================================================================
```

### Package Set-Up and Initialization 

```{r Package Set-Up, include=FALSE}

# Include PHM function libraries
setwd('/d1/home/vleonardo/GPC-Development/bc_qa')  
source('/d1/home/vleonardo/PHM-Development/PHM-LIBRARY.rmd')  # Loads PHM functions
# library(ggplot2)
library(reshape)
PHM.PackageSetup()

# Don't wrap so much
options(width=300)
opts_chunk$set(echo=FALSE)

# BC Function library
source('bc_qa_txform.R')  
```

```{r AddVariable - Add new column to aggregated datamart, based on search strings}

# Note - ignores case and punctuation n the concept path
BCRup.AddVariableToDatamart <- function(p.datamart, p.ptobs, 
                                        p.code.string.1, p.code.string.2, 
                                        p.new.col.name) {
  
  # Sets up return of entire dataset, if this variable raises any exceptions

  tmp.datamart <- p.datamart  
  tmp.exception <- list("")
  tmp.found.and.added.flag <- FALSE
  
  message("Checking variable to be added: ",p.new.col.name) 
  # Ensure only 1 variable matches search terms
  message("... searching variable names containing: '",
          p.code.string.1,"' & '",p.code.string.2,"'")
  tmp.var.found.cnt <- nrow(subset(bs, 
                          grepl(p.code.string.1,variable,ignore.case=TRUE) &
                          grepl(p.code.string.2,variable,ignore.case=TRUE) ))                           
                          
  if (tmp.var.found.cnt > 1) {
    tmp.msg <- "Collision where multiple variables found matching search strings."
    tmp.action.taken <- "WARNING: Variable values not populated."
    tmp.exception <- data.frame(p.new.col.name,p.code.string.1,p.code.string.2,
                           tmp.msg,tmp.action.taken)
    v.exceptions <<- rbind(,v.exceptions, 
                          tmp.exception) } 
  else {
    message("... searching facts for concept paths containing: '",
            p.code.string.1,"' & '",p.code.string.2,"'") 
    tmp.new.col.facts <- subset(p.ptobs, 
                            grepl(p.code.string.1,code.path,ignore.case=TRUE) &
                            grepl(p.code.string.2,code.path,ignore.case=TRUE) )  
    tmp.new.col.facts <- unique(tmp.new.col.facts[,c("patient.num","encounter.num","code","code.descriptor")])
    tmp.new.col.facts <- setNames(tmp.new.col.facts,c("patient.num","encounter.num",p.new.col.name,"code.descriptor"))
    # Ensure observation facts exist for search terms
    tmp.ncf.cnt <- nrow(tmp.new.col.facts)  # Necessary for knitr markdown 
    message("... number of unique rows found: ",tmp.ncf.cnt)
    if (tmp.ncf.cnt == 0) {
      tmp.msg <- "Unable to locate any facts for search strings in concept paths."
      tmp.action.taken <- "WARNING: Variable values not populated."
      tmp.exception <- data.frame(p.new.col.name,p.code.string.1,p.code.string.2,
                             tmp.msg,tmp.action.taken)
      v.exceptions <<- rbind(v.exceptions, 
                            tmp.exception) }
    else { 
      if (tmp.ncf.cnt > nrow(p.datamart)) {
        message("... verifying cardinality of result set for variable requested.")
        tmp.msg <- "Cardinality of variable does not correspond with desired result set."
        tmp.action.taken <- "WARNING: Variable values not populated."
        tmp.exception <- data.frame(p.new.col.name,p.code.string.1,p.code.string.2,
                               tmp.msg,tmp.action.taken)
        v.exceptions <<- rbind(v.exceptions, 
                              tmp.exception) }
      else {
        # All clear, proceed w/ adding variable values to datamart
        tmp.datamart <- merge(tmp.datamart, tmp.new.col.facts,
                              all.x=TRUE)  # don't prune on join mis-matches         
        tmp.found.and.added.flag <- TRUE
        tmp.msg <- "Search terms found."
        message("... number of unique values: ", length(unique(tmp.new.col.facts[,ncol(tmp.new.col.facts)])))
        tmp.action.taken <- "Successfuly added new column!"
        tmp.exception <- data.frame(p.new.col.name,p.code.string.1,p.code.string.2,
                               tmp.msg,tmp.action.taken)
        message(tmp.action.taken)
        v.exceptions <<- rbind(v.exceptions, 
                              tmp.exception)  # Assign value to global variable
        }      
    } 
  }
  
  # If values were not populated, due to wonkiness in search terms, then add NAs instead
  if (! (tmp.found.and.added.flag))  {
     # Go ahead and add placeholder for variable
     #    -- Wonky variables will then just not be populated and instead will be set to NA
     message("Search terms not found....  Column created w/ NA values, instead.")
     tmp.datamart[,c("p.new.col.name")] <- NA
     colnames(tmp.datamart)[colnames(tmp.datamart)=="p.new.col.name"] <- p.new.col.name
     tmp.datamart$code.descriptor <- NA
     message(tmp.action.taken)
     }    
  return(tmp.datamart)
}
```

```{r DescribeVariable - Generate frequency and distribution charts}
# Note - ignores case and punctuation n the concept path
BCRup.DescribeVariable <- function(p.new.col.facts,p.new.col.name) {
 
     # Descriptive analysis for variable added
     tmp.frequency.table <- as.data.frame(addmargins(table(p.new.col.facts[,p.new.col.name])))
     colnames(tmp.frequency.table)[colnames(tmp.frequency.table)=="Var1"] <- p.new.col.name
     colnames(tmp.frequency.table)[colnames(tmp.frequency.table)=="Num of Encounters (Tumors)"] <- p.new.col.name 
  
     print(tmp.frequency.table)  
     print(summary(tmp.frequency.table[,p.new.col.name]))
           
     # Charts
     pie(table(p.new.col.facts[,p.new.col.name]))  # 3 is code
     tmp.frequency.table2 <- aggregate(p.new.col.facts$encounter.num, 
                              by=list(p.new.col.facts$code.descriptor), 
                              function(x) length(unique(x)))
     tmp.frequency.table2 <- tmp.frequency.table2[order(tmp.frequency.table2$x,decreasing=TRUE),]
     tmp.row.cnt <- nrow(tmp.frequency.table2)
     tmp.max.analysis.cnt <- min(7,tmp.row.cnt)
     tmp.chart.subtitle <- "All Occurrences"
     if (tmp.row.cnt > tmp.max.analysis.cnt) {
         tmp.chart.subtitle <- paste0("Most Frequence Occurrences (Top ",tmp.max.analysis.cnt," of ",tmp.row.cnt,")")
     }
     tmp.frequency.table2 <- tmp.frequency.table2[1:tmp.max.analysis.cnt,]  
     op <- par(mar=c(5,4,3,2),bg="white")   # Bottom, left, top, right
     bp <- barplot(tmp.frequency.table2$x,
                     horiz=TRUE,
                     xlim=c(0,max(tmp.frequency.table2$x)*1.25),
                     col=c("lightsteelblue"),
                     #width allows all x labels to be shown
                     xlab="Encounters (Tumors)")         
     text(bp, x=0, labels=tmp.frequency.table2$Group.1, pos=4, cex=.85)
     mtext(side=3, line=1, p.new.col.name, font=2)
     mtext(side=3, line=0, tmp.chart.subtitle) 
     par(op)  # reset
}
```

```{r AddAndDescribeVariable}
BCRup.AddAndDescribeVariableToDatamart <- function(q.datamart, q.ptobs, 
                                        q.code.string.1, q.code.string.2, 
                                        q.new.col.name) {
     q.datamart <- BCRup.AddVariableToDatamart(q.datamart, q.ptobs, 
                                                q.code.string.1,
                                                q.code.string.2,
                                                q.new.col.name) 
     tmp.col.found.flag <- (nrow(q.datamart[! (is.na(q.datamart[,ncol(q.datamart)-1])),]) > 0)
     if  (tmp.col.found.flag) {
        BCRup.DescribeVariable(q.datamart,q.new.col.name) 
        }

     # Remove descriptor f/ dataset to be returned
     q.datamart <- subset(q.datamart,select=-c(code.descriptor))
     
     return(q.datamart)
     }
```

### Loading 'dataset' object

```{r}
#input <- source('dataset.R')$value

# conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, submitter $name, $issues_other
# dataset <- input$about

setwd('/d1/home/vleonardo/GPC-Development/bc-data-files')  
dataset <- list("")
dataset$site <- "KUMC"
# dataset$bc_db <- "bc-redo-alpha-2015-09-08"
# dataset$bc_db <- "bc-redo-subset"
dataset$bc_db <- "KUMC-16-kumcBC"

message('Dataset identified: ',dataset$site)
conn.site <- dbConnect(SQLite(),paste0(dataset$bc_db,".db") )
print(dbListTables(conn.site))

# builder.summary reads raw db
bs <- (builder.summary(conn.site)[, c('variable', 'pat_qty', 'enc_qty', 'fact_qty')])
print(bs)

message('SQLITE db file) patient_dimension: ',
        dbGetQuery(conn.site,"select count(*) from patient_dimension"))
message('SQLITE db file) variable: ',
        dbGetQuery(conn.site,"select count(*) from variable"))
tmp.variable <- dbGetQuery(conn.site,"select * from variable")
print(tmp.variable[,c("concept_path","name_char")])
message('SQLITE db file) concept_dimension: ',
        dbGetQuery(conn.site,"select count(*) from concept_dimension"))
message('SQLITE db file) observation_fact: ',
        dbGetQuery(conn.site,"select count(*) from observation_fact"))

```

### Load and Merge Observations With Concept Descriptors

``` {r}
############################################################################
# Delaware: n=4660583
# Expected Run-time: 6.5 mins
ptobs <- PHM.LoadObservationsWithConceptDescriptors(conn.site)

saveRDS(ptobs,  paste0(dataset$bc_db,"-",Sys.Date(),"-pt-obs.rds"))
print(nrow(ptobs))
ptobs.code.paths <- unique(ptobs$code.path)                           
```

### Loading original patient data for `r dataset$site'
 -- variables used in QA process (exclusion criteria and receptor status)

```{r}
setwd('/d1/home/vleonardo/GPC-Development/bc_qa')  
load("bc_terms_results.RData")
message('Number of terms loaded: ',nrow(bcterm$term204))

message(Sys.time())
tumor.site <- bc.exclusions(conn.site)   # Expected run-time: 10 mins for 14k pts
message(Sys.time())

print(matrix(names(tumor.site)))
saveRDS(tumor.site,  paste0(dataset$bc_db,"-",Sys.Date(),"-tumor-site.rds"))
x.tumor.site <- tumor.site  # Holding area for restoration/debugging purposes
#tumor.site <- readRDS(paste0(dataset$bc_db,"-",Sys.Date(),"-tumor-site.rds"))
message('Number of tumors loaded: ',(nrow(tumor.site)))
message('Number of pts in dataset: ',length(unique(tumor.site$patient_num)))


survey.sample <- check.cases(tumor.site)  # Creates tumor-level criteria flags
survey.sample$all.criteria <- reduce.logical(survey.sample)
message('Number of pts meeting inclusion criteria: ', 
        nrow(survey.sample[survey.sample$all.criteria,]))

survey.sample.size <- count.cases(survey.sample)  # Produce table of criteria counts
print(survey.sample.size)
```

### Filter for only eligible encounters(tumors)
```{r}
message("Total number of patients submitted: ", length(unique(tumor.site$patient_num)))
message("Total number of encounters submitted: ", length(unique(tumor.site$patient_num)))
tumor.site <- subset(tumor.site,encounter_num %in% survey.sample$encounter_num[survey.sample$all.criteria])
message("Number of eligible patients: ", length(unique(tumor.site$patient_num)))
message("Number of eligible encounters(tumors): ", nrow(tumor.site))

```

### Rename for R syntax

```{r}
colnames(tumor.site)[colnames(tumor.site)=="patient_num"] <- "patient.num"
colnames(tumor.site)[colnames(tumor.site)=="encounter_num"] <- "encounter.num"
```

### Report only the first eligible encounter for each patient

```{r}
#message('Number of pts w/ multiple eligible encounters: ',
        #length(tumor.site[duplicated(tumor.site$patient.num),]$patient.num) )
#tmp.first.enctrs <- setNames(
#                       aggregate(tumor.site$encounter.num, by=list(tumor.site$patient.num),min),
#                       c("patient.num","encounter.num") )
#tumor.site <- merge(tumor.site,tmp.first.enctrs,by=c("patient.num","encounter.num"))
#message('Number of pts w/ multiple eligible encounters: ',
#        length(tumor.site[duplicated(tumor.site$patient.num),]$patient.num) )
```

### Trim dataset for only eligible encounters

```{r}
ptobs <- subset(ptobs, encounter.num %in% tumor.site$encounter.num)
print(nrow(ptobs))
```


### Add next variables, specifying 2 search terms for 'variable name'

```{r}
# Variable Initialization
v.exceptions <- data.frame(p.new.col.name="",p.code.string.1="",p.code.string.2="",tmp.msg="",tmp.action.taken="")
message("Rows in tumor.site: ", nrow(tumor.site))
tmp.tumor.site <- tumor.site
tmp.tumor.site <- cbind(gpc.site=dataset$site,tmp.tumor.site)
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'Breast','Breast', 'Seer.Site.Breast')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0160','race 1', 'NAACCR0160.Race.1')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0161','race 2', 'NAACCR0161.Race.2')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0162','race 3', 'NAACCR0162.Race.3')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0163','race 4', 'NAACCR0163.Race.4')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0164','race 5', 'NAACCR0164.Race.5')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0190','spanish', 'NAACCR0190.Spanish')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0220','sex', 'NAACCR0220.Sex')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0240','birth', 'NAACCR0240.DOB')  # Date of birth
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0380','central', 'NAACCR0380.Seqno.Central')  # Seqno Central
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0390','diagnosis', 'NAACCR0390.Dx.Date')  # Date of diagnosis
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0400','primary', 'NAACCR0400.Primary.Site')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0410','laterality', 'NAACCR0410.Laterality')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0440','grade', 'NAACCR0440.Grade')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0490','confirmation', 'NAACCR0490.Confirmation')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0560','hospital', 'NAACCR0560-Seqno.Hosp')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0610','class', 'NAACCR0610-Class.Case')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0670','surg', 'NAACCR0560-Surg.Prim.Site')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0820','positive', 'NAACCR0820.Reg.Nodes.Pos')  # Regional Nodes Positive
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'0830','examine', 'NAACCR0820.Reg.Nodes.Examined')  # Regional Nodes Examined
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'1750','contact', 'NAACCR0820.Date.Last.Contact')  
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'1760','vital', 'NAACCR1760.Vital.Status')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'1860','recurrence', 'NAACCR1860.Recurrence.Date')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'1861','flag', 'NAACCR1860.Recurrence.Date.1st.Flag')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2850','dx', 'NAACCR2850.CSMets.Dx')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2860','eval', 'NAACCR2860.CSMets.Eval')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2869','factor', 'NAACCR2869.CSSF.15')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2876','factor', 'NAACCR2876.CSSF.22')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2877','factor', 'NAACCR2877.CSSF.23')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2880','factor', 'NAACCR2880.CSSF.01')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2890','factor', 'NAACCR2890.CSSF.02')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'2940','AJCC-6', 'NAACCR2940.AJCC-6.T')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'3000','AJCC-6', 'NAACCR3000.AJCC-6.Stage')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'3020','SS2000', 'NAACCR3020.SS2000')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'3400','AJCC-7', 'NAACCR3400.AJCC-7.T')
tmp.tumor.site <- BCRup.AddAndDescribeVariableToDatamart(tmp.tumor.site, ptobs,'3430','AJCC-7', 'NAACCR3430.AJCC-7.Stage')

names(tmp.tumor.site)
message("Rows in tumor.site: ", nrow(tumor.site))
print(v.exceptions)
```


### Export dataset to CSV file

```{r Send patient sampling to CSV file}
setwd('/d1/home/vleonardo/GPC-Development/bc-datamart')  

# Remove original recept status variables f/ inclusion set
csv.tumor.site <- subset(tmp.tumor.site,select=-c(er.csf.1,pr.csf.2,her2.csf.15,mgs.method.csf.22,mgs.score.csf.23))

message(paste('Output directory: ',getwd()) )
v.filename <- paste0(dataset$site,"-",dataset$record_id,"-BCDatamart.csv")
write.csv(csv.tumor.site, file = v.filename, row.names=FALSE)
message('Tumor-level datamart: ',v.filename)
```