Tidy data from i2b2
===================

The [i2b2 star schema][fig4] is not [tidy][]:
  - the fact table is a long-skinny
  - the concept and modifier dimensions are heirarchical

[fig4]: http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3000779/figure/fig4/
[tidy]: https://www.jstatsoft.org/article/view/v059i10


```{r libs_opts, echo=FALSE}
library(reshape2)
options(width=180)

source('i2b2_shape.R')
```

## i2b2 data submitted from a GPC site

```{r input, echo=FALSE}
input <- source('dataset.R')$value

conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, ...
dataset <- input$about
dataset[, c('record_id', 'site', 'name',
            'bc_db', 'content_length', 'timestamp')]
```

## Tidy data for Patients

This is pretty straightforward...

```{r}
pat <- dbGetQuery(conn.site, 'select * from patient_dimension')
head(pat[-19:-11])
```

**TODO**: consider factors for `..._cd` columns.


## Data dictionary: NAACCR standard terminology and GPC Query Terms

Recall the i2b2 query terms relevant to inclusion/exclusion criteria for
the [GPC breast cancer survey study][GPC_BC]:

[GPC_BC]: http://www.public-health.uiowa.edu/herce/research/gpc/index.html

```{r}
load("bc_terms_results.RData")
# names(bcterm)
t.incl <- bcterm$t.incl
t.excl <- bcterm$excl.all

items <- rbind(t.incl, t.excl)
# Items without paths are synthetic combinations of other items.
items <- subset(items, !is.na(items$concept_path))
# strip trailing whitespace. odd.
items$name <- gsub(' +$', '', items$name)

items
```

Applying the `data.dictionary` to our SQL data gives us convenient access
to terminology in our dataset; for example, the names corresponding to paths:

```{r}
dd <- data.dictionary(conn.site)
head(dd$path.name)
```

The `norm.path` function prunes some irrelevant info and makes an R factor:
```{r}
items$concept_path <- dd$norm.path(items$concept_path)
```

## Nominal per-tumor data: Primary Site

Recall each tumor record from NAACCR has a distinct i2b2 `encounter_num`.

The `per.encounter.nominal` function gives us all the nominal data on
a per-encounter basis.

For example, primary site is one of the terms from our query; for each
encounter: we see the code, name, and path of the primary site:

```{r}
enc.nom <- per.encounter.nominal(conn.site, dd$code.levels)
head(enc.nom)

primary.site <- dd$code.name(dd$obs.under(enc.nom, t.incl['primary.site', 'concept_path']))

# Merge puts the join key first; put the encounter_num back where it was.
primary.site <- primary.site[, c('encounter_num', 'concept_cd', 'name_char', 'concept_path')]
head(primary.site)
```

## Multiple variables: long and wide formats

The `nominal.log` function gives us data on several items in long format:

```{r}
long <- nominal.long(enc.nom, dd, items)
unique(long$variable)
head(long)
```

We can then use conventional R tools such as `dcast` from [reshape][].

[reshape]: https://github.com/hadley/reshape

```{r}
wide.coded <- dcast(long, encounter_num ~ variable, value.var='concept_cd')
tail(wide.coded)
wide.labelled <- dcast(long, encounter_num ~ variable, value.var='name_char')
tail(wide.labelled)
```

## Dates

```{r}
qterms <- bcterm$term204
qterms$concept_path <- dd$norm.path(qterms$concept_path)

event.vars <- qterms[grepl('Date', qterms$name)
                     | grepl('0400 Primary Site', qterms$name)
                     | grepl('SEER Site\\\\Breast', qterms$concept_path), ]
rownames(event.vars) <- 
  gsub(' $', '', gsub('Date', '', gsub('Date of ', '', event.vars$name)))

event.vars
```


```{r}
when <- per.encounter.date(conn.site, dd$code.levels)
head(when)

head(dd$obs.at(when, event.vars[1,]$concept_path))
```

```{r}
when.long <- date.long(when, dd, event.vars)
head(when.long)

# We don't seem to have values for all of these date variables.
unique(when.long$name_char)

when.wide.start <- dcast(when.long,
                       encounter_num ~ variable, value.var='start_date',
                       drop=FALSE)

# BLARGH!
# http://stackoverflow.com/questions/12289731/posixct-values-become-numeric-in-reshape2-dcast
when.wide.start[-1] <- lapply(when.wide.start[-1], as.POSIXct, origin="1970-01-01")
head(when.wide.start)

when.wide.end <- dcast(when.long,
                       encounter_num ~ variable, value.var='end_date',
                       drop=FALSE)
when.wide.end[-1] <- lapply(when.wide.end[-1], as.POSIXct, origin="1970-01-01")
head(when.wide.end)
```

## Colophon: Version Control

```{r version, echo=FALSE}
version.bc_qa <- try (
  version.bc_qa <- paste(' version', system(' git rev-parse HEAD', intern=TRUE)),
  silent=TRUE
  )
```

source: [bc_qa](https://github.com/kumc-bmi/bc_qa) `r version.bc_qa`
