```{r libraries, echo=FALSE}
library(ggplot2)
library(reshape)

source('bc_qa_txform.R')

```


```{r wrap, echo=FALSE}
# Don't wrap so much
options(width=300)
```


```{r site.input, echo=FALSE}
input <- source('dataset.R')$value

conn.site <- input$conn

# about has $record_id $site name, $bc_db filename, $content_length, submitter $name, $issues_other
dataset <- input$about
```

```{r terms, echo=FALSE}
load("bc_terms_results.RData")
```

Multi-tumor patients at `r dataset$site`
========================================

While looking at sequence number, let's consider patients who have multiple otherwise-qualifying tumors.
We show such cases left-to-right in order of their earliest diagnosis, with a segment connecting the first
and last date of (qualifying) diagnosis. The color of the dot at the recent end of the segment indicates
its sequence number (if available?). Note that in several cases, the tumors are diagnosed on the same day,
so the choice of which tumor to use to color the dot is arbitrary.

```{r}
tumor.site <- bc.exclusions(conn.site)
survey.sample <- check.cases(tumor.site)
```


```{r}
but.prior <- subset(survey.sample,
                    select=-c(encounter_num, patient_num, age, span, no.prior
                              # tried skipping demographics to look at MCRF data...
                              # adult, female, not.dead, english
                              ))
maybe.prior <- reduce.logical(but.prior)

tumor.per.pat <- data.frame(table(tumor.site$patient_num[maybe.prior]))

focus <- subset(tumor.site,
       patient_num %in% subset(tumor.per.pat, Freq > 1)$Var1 &
         maybe.prior,
       select=c(encounter_num, patient_num, date.dx, seq.no, primary.site))
```

```{r}
x <- unique(dx.span(focus)[, c('patient_num', 'first', 'last', 'span')])

x <- merge(x, focus[, c('patient_num', 'date.dx', 'seq.no')],
           by.x=c('patient_num', 'first'), by.y=c('patient_num', 'date.dx'))

if (!any(!is.na(x$seq.no))) { x$seq.no <- '?' } # ggplot barfs if all fill=seq.no are NA

x <- x[order(x$first), ]
x$ix <- 1:nrow(x)
```

```{r seq.dates, echo=FALSE, fig.width=12, fig.height=8}
ggplot(x, aes(x=ix, xend=ix, y=first, yend=last, fill=seq.no)) +
  geom_segment() +
  geom_point(aes(x=ix, y=first, color=seq.no)) +
#  geom_bar(aes(x=ix, y=span), stat="identity") +
#  coord_cartesian(ylim=c(min(x$first), max(x$last))) +
  labs(x="Otherwise Qualified Cases", y="time",
       title="First and last diagnosis date")
```

Details of these cases:

```{r}
focus
```
